% ===================================================================
% Appendix D — Golden Mirror
% Stub scaffold (R6 lane discipline). Body to be filled by L17/L23/L27
% lane authors who own the Trinity-vs-Pellis comparison.
% Refs: gHashTag/trios#101 · #30 · #63
% Source of truth (read-only from this lane):
%   - trinity-clara/proofs/igla/lucas_closure_gf16.v
%   - trinity-clara/proofs/pellis/  (Pellis embedding theorems)
%   - cargo run -p tri-math --bin tri-math -- math compare --pellis
%   - Zenodo DOI 10.5281/zenodo.19227879 (Pellis embedding)
% R5 honesty: this stub does not assert numerical equality between
% Trinity and Pellis; it only fixes the comparison axis so the
% body-fill lane can write down honest agreements and disagreements.
% ===================================================================

\chapter{Golden Mirror}
\label{app:golden-mirror}

This appendix records the side-by-side comparison of the Trinity
identity ($\varphi^{2} + \varphi^{-2} = 3$) and the Pellis embedding
(Zenodo DOI \href{https://doi.org/10.5281/zenodo.19227879}%
{10.5281/zenodo.19227879}) on three axes: (i) algebraic outputs of
the two formal frameworks on identical inputs, (ii) the hybrid
inner-product calculation that bridges the two metrics, and (iii)
the ultraviolet / infrared (UV/IR) renormalisation-group (RG)
analysis under each metric.  The appendix is the empirical mirror
of the theoretical embedding worked out in
Chapter~\ref{ch:trinity-identity} and recalled in the related-work
Chapter~\ref{ch:related}.

\section{Comparison axes and source of truth}

The numerical anchor of every row in this appendix is the binary
\verb|cargo run -p tri-math -- math compare --pellis|, whose output
is pinned to the commit recorded alongside the row in
\verb|assertions/seed_results.jsonl| (Pellis-mirror rows). The
Coq source for the Trinity side lives in
\verb|trinity-clara/proofs/igla/lucas_closure_gf16.v| (theorem
\verb|lucas_2_eq_3| --- Proven, INV-5 base case), and the Pellis
side is anchored at the Pellis-embedding repository at the Zenodo
DOI cited above.

% TODO[L17,L23,L27]: replace this paragraph with the audited
% comparison sentence: ``On the test inputs listed in Table D.1,
% the Trinity and Pellis frameworks agree to <X> decimal digits on
% the algebraic axis and to <Y> decimal digits on the hybrid axis.''
% Honest blanks until the comparison binary produces real output.

\section{Side-by-side: Trinity vs.\ Pellis}
\label{sec:appD-side-by-side}

% TODO[L27]: insert Table D.1 with columns
% [input expression, Trinity output, Pellis output, $\Delta$,
%  agreement status, source commit].
% Each row's source commit must be pinned at the time of the run;
% any seven-digit SHA on the embargoed list is auto-rejected (R5).

\subsection{Falsifier}
Any single row of Table~\ref{sec:appD-side-by-side} on which the
Trinity and Pellis outputs disagree beyond the published numerical
tolerance falsifies the embedding claim of
Chapter~\ref{ch:trinity-identity} for that input regime, and the
chapter must be revised or its scope narrowed.

\section{Hybrid inner product}
\label{sec:appD-hybrid}

% TODO[L23]: write the explicit form of the hybrid inner product
% \langle u, v \rangle_h = \alpha \langle u, v \rangle_T
%                       + (1 - \alpha) \langle u, v \rangle_P
% where \alpha \in [0,1] interpolates between the Trinity (T) and
% Pellis (P) inner products. Acceptance criterion: positive-definite
% on the test basis listed in trinity-clara/benchmarks/hybrid_basis.csv.

\subsection{Falsifier}
A test vector for which the hybrid inner product fails to be
positive-definite at any $\alpha \in [0,1]$ falsifies the embedding
hypothesis and forces a metric correction.

\section{UV/IR RG analysis}
\label{sec:appD-rg}

% TODO[L17,L23]: insert Figure D.1 --- RG flow diagram showing UV
% and IR fixed points under the Trinity metric (left panel) and the
% Pellis metric (right panel), overlaid by the hybrid metric flow
% (centre panel).  Cite ch:trinity-identity and the related-work
% Chapter~\ref{ch:related} survey of RG-anchored sacred-geometry
% interpretations.

\subsection{Falsifier}
A UV or IR fixed point that exists under the Trinity metric but
fails to exist under the hybrid metric for some
$\alpha \in (0,1)$ falsifies the smooth-interpolation claim and
requires either (i) a piecewise-defined hybrid metric, or (ii) a
narrower domain for the embedding.

\section{Audit semantics}

The audit utility
\verb|cargo run -p trios-phd -- audit --pellis| iterates over the
three tables / figures of this appendix and verifies that every row
references a SHA pinned in
\verb|assertions/seed_results.jsonl| (Pellis-mirror category) and
not on the embargoed SHA list. Any unpinned numeric claim halts the
build with a non-zero exit code.

\section{Reading this appendix honestly (R5)}

The Trinity / Pellis comparison is, as of the cycle in which this
scaffold was written, \emph{open}: no honest comparison rows exist
in \verb|assertions/seed_results.jsonl|. The body-fill lane must
populate the tables only after the comparison binary has produced
real output on a non-embargoed commit; the present scaffold
deliberately contains no fabricated rows (R5).

% TODO[L17,L23,L27,LP]: once Appendix~\ref{app:falsification} is
% extended with the Pellis-mirror falsifiers, add a back-reference
% from each falsifier subsection above to the corresponding row of
% Table~\ref{tab:falsification}.

%% End of Appendix D scaffold.
%% Hand-off note: the next agent to claim this appendix should post on
%% trios#265 with `🔒 AGENT <id> CLAIMING: App.D body-fill --- L17/L23/L27 mirror`
%% and follow phd-chapter-author SKILL.md Steps 2-9.
