% ===================================================================
% Appendix H — ACM Artifact Evaluation Checklist
% Targets: Functional, Reusable, Available (3-badge submission).
% ===================================================================

\chapter{ACM Artifact Evaluation Checklist}
\label{app:acm-ae}

This appendix records the artefact-evaluation submission against the
ACM Artifact Review and Badging policy. The submission targets the
three available badges.

\section{Artifact summary}

\begin{description}
  \item[Title.] \emph{Flos Aureus} — Trinity Framework artefact pack.
  \item[Authors.] Dmitrii Vasilev (Trinity Research Programme).
  \item[Repository.] \url{https://github.com/gHashTag/trios}, branch
        tagged \verb|phd/v1.0| at submission.
  \item[License.] MIT (code), CC-BY-4.0 (text), Apache-2.0 (Coq proofs).
\end{description}

\section{Functional badge}

The \emph{Functional} badge is awarded to artefacts that are
documented, complete, exercisable, and consistent with the claims in
the paper.

\begin{itemize}
  \item Documentation entry point: \texttt{docs/phd/reproducibility.md}.
  \item Build entry point: \verb|cargo run -p trios-phd -- compile|.
  \item Tests exercised by reviewer: \verb|cargo test -p trios-phd|
        and \verb|cargo test -p trios-igla-race -- invariants|.
  \item Smoke run: \verb|cargo run -p trios-phd -- reproduce --chapter 24|
        recovers Table~24.1 within \(\pm 0.5\%\) on seeds
        \(\{17, 42, 1729\}\).
\end{itemize}

\section{Reusable badge}

The \emph{Reusable} badge requires that the artefact can be reused by
others, with documentation that supports adaptation.

\begin{itemize}
  \item All numeric constants traceable to a \verb|.v| file via the
        single source of truth (rule L-R14).
  \item Reviewer-facing manifest at \texttt{docs/phd/reproducibility.md}
        listing dependencies, hardware profile, expected wall-clock,
        and SHA-256 checksums for every CSV.
  \item Modular structure: per-chapter reproduction is selectable via
        \verb|--chapter NN|.
  \item Public API: the runtime guard layer
        (\texttt{crates/trios-igla-race}) exports
        \verb|enforce_all_invariants(...)| as a stable entry point
        for downstream reuse.
\end{itemize}

\section{Available badge}

The \emph{Available} badge requires permanent retrieval of the
artefact via a stable identifier.

\begin{itemize}
  \item Zenodo deposit (DOI in Appendix~G).
  \item GitHub source mirror at the \verb|phd/v1.0| tag.
  \item License files (\texttt{LICENSE-MIT}, \texttt{LICENSE-CC-BY},
        \texttt{LICENSE-APACHE-2.0}) shipped at repository root.
\end{itemize}

\section{Reviewer instructions (compact)}

\begin{enumerate}
  \item Clone the tag: \\
        \verb|git clone --branch phd/v1.0 https://github.com/gHashTag/trios|.
  \item Install Rust toolchain (stable, version pinned in
        \verb|rust-toolchain.toml|).
  \item Run the audit gate: \verb|cargo run -p trios-phd -- audit|.
  \item Reproduce a chapter: \verb|cargo run -p trios-phd -- reproduce --chapter 24|.
  \item Verify hashes against \verb|docs/phd/reproducibility.md|.
\end{enumerate}

\noindent
A reviewer who hits any deviation greater than \(\pm 0.5\%\) on a
reported table is invited to file an issue at
\href{https://github.com/gHashTag/trios/issues}{the issue tracker}; the
reproduction binary's deterministic seeding makes this self-debugging.
