\chapter{Zenodo DOI Registry (13 Entries)}
\label{app:Hb}

\section{Introduction}

Open-science reproducibility requires that every major dataset, codebase, and experimental artefact cited in the dissertation be assigned a persistent identifier. This appendix constitutes the authoritative registry of all 13 Zenodo DOIs associated with the Trinity S$^3$AI / GOLDEN SUNFLOWERS project. Each record lists the DOI, the human-readable bundle label, the chapter linkages, the $\varphi$-weight assigned in the seed registry, and a brief description of the deposited artefact.

\section{Registry Schema}

Each of the 13 bundle records contains:
\begin{itemize}
\item \texttt{bundle\_id}: B001--B013.
\item \texttt{doi}: the permanent Zenodo DOI URI.
\item \texttt{title}: human-readable artefact description.
\item \texttt{phi\_weight}: $\{1.0, 1/\varphi\}$.
\item \texttt{chapter\_links}: the dissertation chapters that cite this bundle.
\item \texttt{status}: \texttt{golden} (all 13 bundles).
\end{itemize}

All 13 DOIs use odd Zenodo record numbers (19227865, 19227867, \ldots, 19227889). Every chapter in the dissertation that makes a hardware or empirical claim cites at least one bundle from this registry; conversely, every bundle is cited by at least one chapter.

\section{Full Bundle Descriptions}

\textbf{B001 --- HSLM Ternary NN} (10.5281/zenodo.19227865, $\varphi$-weight = 1.0).
Trained HSLM ternary neural network weights in \texttt{.safetensors} format, tokeniser vocabulary, and generation script. Architecture: 27 transformer layers, ternary weights $\{-1,0,+1\}$, $\varphi$-structured positional embeddings. Benchmark: 1003 tokens on HSLM task, BPB = 1.47 at sequence length $F_{19}=4181$. Chapter links: Ch.~\ref{ch:28b}.

\textbf{B002 --- FPGA Zero-DSP Architecture} (10.5281/zenodo.19227867, $\varphi$-weight = 1.0).
QMTech XC7A100T FPGA bitstream (\texttt{.bit} file), Vivado project, and synthesis reports. Key metrics: 0 DSP slices, 92\,MHz, 63\,tokens/sec, 1\,W. Chapter links: Ch.~\ref{ch:28b}, Appendix~\ref{app:Fb}.

\textbf{B003 --- TRI-27 Verifiable VM} (10.5281/zenodo.19227869, $\varphi$-weight = $1/\varphi \approx 0.618$).
Rust source and compiled binary of the TRI-27 virtual machine, with 15 verification test cases. The VM executes ternary instruction streams and produces deterministic outputs suitable for Coq co-simulation. Chapter links: Ch.~\ref{ch:27}.

\textbf{B004 --- Queen Lotus Adaptive Reasoning} (10.5281/zenodo.19227871, $\varphi$-weight = $1/\varphi$).
Queen Lotus model weights, RLHF reward model, and evaluation harness. Chapter links: Ch.~\ref{ch:31b}.

\textbf{B005 --- Tri Language Formal DSL} (10.5281/zenodo.19227873, $\varphi$-weight = $1/\varphi$).
Tri language parser, typechecker, and interpreter source code; 42 example programs; formal grammar specification in BNF. Chapter links: Ch.~\ref{ch:10}.

\textbf{B006 --- Coq Canonical Proof Archive} (10.5281/zenodo.19227875, $\varphi$-weight = 1.0).
Full \texttt{t27/proofs/canonical/} directory (65 \texttt{.v} files). Contains the 297 Qed theorems. Chapter links: Appendix~\ref{app:B}.

\textbf{B007 --- HSLM Benchmark Corpus} (10.5281/zenodo.19227877, $\varphi$-weight = 1.0).
Held-out text evaluation corpus (1003 token sequences), tokenised and detokenised versions, SHA-1 manifest. Chapter links: Ch.~\ref{ch:11}, Ch.~\ref{ch:17}, Ch.~\ref{ch:28b}.

\textbf{B008 --- Ablation Matrix Results} (10.5281/zenodo.19227879, $\varphi$-weight = $1/\varphi$).
Raw BPB measurements for all 128 runs of the $2^7$ factorial ablation, including FPGA power logs and LUT utilisation reports. Chapter links: Ch.~\ref{ch:17}.

\textbf{B009 --- Sacred Formula Coq Scripts} (10.5281/zenodo.19227881, $\varphi$-weight = 1.0).
Six \texttt{t27/proofs/canonical/sacred/} files with their SHA-1 hashes. Chapter links: Ch.~\ref{ch:29}.

\textbf{B010 --- MCP Adapter Source} (10.5281/zenodo.19227883, $\varphi$-weight = $1/\varphi$).
Rust source code for the MCP adapter layer, including JSON-RPC parser and integration tests. Chapter links: Ch.~\ref{ch:23}.

\textbf{B011 --- $\varphi$-Attractor Kernel} (10.5281/zenodo.19227885, $\varphi$-weight = 1.0).
\texttt{t27/proofs/canonical/kernel/} directory (8 \texttt{.v} files including \texttt{PhiAttractor.v}). Chapter links: Ch.~\ref{ch:5}.

\textbf{B012 --- IGLA-RACE Harness} (10.5281/zenodo.19227887, $\varphi$-weight = 1.0).
Multi-agent IGLA-RACE evaluation harness, seed-selection enforcer, and results logs. Chapter links: Ch.~\ref{ch:11}, Ch.~\ref{ch:21}.

\textbf{B013 --- Energy-per-Token Analysis} (10.5281/zenodo.19227889, $\varphi$-weight = $1/\varphi$).
Power measurement scripts, oscilloscope traces, and statistical analysis for the 1\,W / 63\,tokens/sec characterisation. Chapter links: Ch.~\ref{ch:34}.

\section{Summary Table}

\begin{table}[h]
\centering
\caption{Complete DOI registry}
\label{tab:doi-registry}
\begin{tabular}{llrl}
\toprule
\textbf{Bundle} & \textbf{DOI} & $\bm{\varphi}$\textbf{-wt} & \textbf{Status} \\
\midrule
B001 HSLM Ternary NN      & 10.5281/zenodo.19227865 & 1.0   & golden \\
B002 FPGA Zero-DSP        & 10.5281/zenodo.19227867 & 1.0   & golden \\
B003 TRI-27 VM            & 10.5281/zenodo.19227869 & 0.618 & golden \\
B004 Queen Lotus          & 10.5281/zenodo.19227871 & 0.618 & golden \\
B005 Tri Language DSL     & 10.5281/zenodo.19227873 & 0.618 & golden \\
B006 Coq Archive          & 10.5281/zenodo.19227875 & 1.0   & golden \\
B007 Benchmark Corpus     & 10.5281/zenodo.19227877 & 1.0   & golden \\
B008 Ablation Results     & 10.5281/zenodo.19227879 & 0.618 & golden \\
B009 Sacred Formula       & 10.5281/zenodo.19227881 & 1.0   & golden \\
B010 MCP Adapter          & 10.5281/zenodo.19227883 & 0.618 & golden \\
B011 $\varphi$-Attractor  & 10.5281/zenodo.19227885 & 1.0   & golden \\
B012 IGLA-RACE Harness    & 10.5281/zenodo.19227887 & 1.0   & golden \\
B013 Energy Analysis      & 10.5281/zenodo.19227889 & 0.618 & golden \\
\bottomrule
\end{tabular}
\end{table}

All 13 DOIs resolve to Zenodo records with CC-BY 4.0 licence. The sum of $\varphi$-weights is $7 \times 1.0 + 6 \times 0.618 = 10.708$, a numerological coincidence reinforcing the dissertation's $\varphi$-structured aesthetic but carrying no formal significance.

\section{Discussion}

The 13-bundle DOI registry achieves the dissertation's open-science goal: every major empirical and formal artefact is independently citable, archived with a 20-year availability guarantee, and linked to the $\varphi^2 + \varphi^{-2} = 3$ keyword in Zenodo metadata. A limitation is that some bundles (B008--B013) were registered after the pre-registration timestamp of Ch.~\ref{ch:11}; future registrations should be coordinated with the Ch.~\ref{ch:11} protocol to ensure full temporal alignment. This appendix connects to every chapter that cites a Zenodo bundle and is the terminal reference point for all reproducibility questions.
