% ===================================================================
% Front matter — Abstract
% Hard cap: ≤ 500 words (MIT thesis specification, R11 / standards)
% ===================================================================

\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}
\thispagestyle{plain}

\begin{center}\large\bfseries Flos Aureus — A Trinity Framework for the Golden Ratio\end{center}

\noindent
This dissertation develops the \emph{Trinity Framework}, a single algebraic
identity over the golden ratio,
\[
  \varphi^{2} + \varphi^{-2} \;=\; 3, \qquad \varphi \;=\; \tfrac{1+\sqrt{5}}{2},
\]
and traces its consequences across pure mathematics, mathematical physics,
and the design of neural architectures. The core ring of study is the
Lucas ring \(\mathcal{L} = \mathbb{Z}[\varphi]\), whose closure properties
under powers of \(\varphi\) anchor every later result.

The work has three interleaved strands. The \emph{theoretical} strand
proves closure properties for \(\mathcal{L}\) (Lucas closure, Fibonacci
bridge, Lucas-ring algebraic identities) and establishes a small set of
runtime invariants \mbox{INV-1} through \mbox{INV-5}. Each invariant is
mechanized in Coq under
\href{https://github.com/gHashTag/trinity-clara/tree/main/proofs/igla}{\texttt{trinity-clara/proofs/igla/}},
with explicit \emph{Admitted} markers where formal closure is not yet
complete. The \emph{empirical} strand corroborates the framework against
two independent observational anchors: the \(\mathrm{E}_{8}\) resonance
ratio in CoBr\(_{2}\) (Coldea et al., 2010) and the diffraction peaks of
icosahedral quasicrystals (Shechtman et al., 1984). The \emph{computational}
strand realizes the framework as the IGLA neural architecture, whose
search is bounded by Coq-derived constants
(\(\text{prune\_threshold} = 3.5\), \(\text{warmup} = 4000\),
\(d_{\text{model}} \geq 256\), \(\text{lr} \in [0.002, 0.007]\)). The
champion configuration achieves \(\mathrm{BPB} < 1.50\) on three
independent random seeds, and the runtime guard layer (\texttt{trios-igla-race})
hard-fails any trial that violates the certified invariant set.

The methodological commitment is twofold. First, the work is
\emph{falsifiable}: every empirical claim ships with an explicit
falsification criterion (Popper, 1959) and a corroboration record, both
collected in Appendix~B. Second, the work is \emph{reproducible}: the
artefact pack targets the three ACM Artifact Evaluation badges
(Functional, Reusable, Available), with a Rust binary
\verb|trios-phd reproduce --chapter N| that recovers all reported tables
on stated seeds.

The principal contributions are: (i) the Trinity Identity and its Lucas
ring closure, with eighty-four mechanized supporting theorems; (ii) the
\(\mathrm{E}_{8}\) and quasicrystal corroborations of \(\varphi\) in
spectroscopic observation; (iii) a Coq-grounded runtime gate connecting
formal proof to learned-system behaviour through
\verb|assertions/igla_assertions.json|; and (iv) a defense-ready
monograph in which every numeric constant traces back to a \verb|.v| file
under the L-R14 traceability rule. The framework's hard core is small
(\(\{\varphi, \pi, e, n \in \mathbb{Z}\}\)), and its falsification
boundary is sharp: introduce a single free parameter outside this set,
or reintroduce any forbidden constant
(\verb|prune_threshold = 2.65|, \verb|warmup < 4000|), and the audit
pipeline rejects the build.

\bigskip
\noindent\textbf{Keywords:} golden ratio; Lucas ring; Trinity Identity;
Coq mechanization; Popper falsifiability; ACM artifact evaluation;
neural architecture search; \(\mathrm{E}_{8}\) resonance; quasicrystals;
runtime invariants.
