% ===================================================================
% Front matter — Acknowledgements
% Cambridge / Springer LNM convention.
% Monograph: Flos Aureus: A Theorem-Guided Architecture Search
%            through the Trinity Identity φ²+φ⁻²=3
% Author:    Кирилл Шамин (Cyril Shamin)
% Institute: University of Reading / gHashTag Open Doctorate
% ===================================================================
%
% TODO: This is a scaffold stub. The text below provides section
%       headings and placeholder prose. Replace with the author's
%       personal acknowledgements before submission:
%
%   TODO-1 (Family): Add personal dedication to family members.
%           Cambridge §3.4 encourages naming family support.
%   TODO-2 (Supervisor): Name the primary supervisor(s), their
%           institution, and the nature of their guidance.
%           Confirm supervisor consent before naming them.
%   TODO-3 (Trinity Hive Agents): Decide whether autonomous agents
%           (LF, LB, LC, … lanes) are acknowledged here or in the
%           Preface. University of Reading AI-use policy may
%           require explicit disclosure — check and comply.
%   TODO-4 (Funding): List all funding sources with grant numbers.
%           Required by most universities (Cambridge §3.4,
%           UKRI grant acknowledgement wording).
%   TODO-5: Remove the \subsection*{} headings if the institution
%           requires a single unstructured acknowledgements page.
% ===================================================================

\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}
\thispagestyle{plain}

This monograph stands on three shoulders that the author wishes to
name explicitly.

The first is the long lineage of mathematicians who insisted that
\(\varphi\) is not an aesthetic ornament but a structural principle:
from Euclid's \emph{Elements} (Book~II, prop.~11) and Pacioli's
\emph{De Divina Proportione} (1509) to Lucas's number sequence (1878),
Penrose's tilings (1974), and Shechtman's quasicrystal observation
(1982). Their work made it intellectually safe to take \(\varphi\)
seriously as physics.

The second is the community of formal-methods practitioners whose work
made it technically possible to mechanize this dissertation in Coq:
the \texttt{coqc} maintainers, the authors of \texttt{Coq.Interval}
and the standard library, and in particular the broader ITP community
whose insistence on \emph{Admitted} honesty (R5) shaped the editorial
discipline of every chapter that follows.

The third is the experimental physics community whose corroboration
records gave the framework an empirical falsifier: R.~Coldea and
collaborators for the \(\mathrm{E}_{8}\) resonance work in
CoBr\(_{2}\) (\emph{Science}, 2010), and the long quasicrystal
literature beginning with Shechtman \emph{et al.} (1984).

% -------------------------------------------------------------------
\subsection*{Family}
% TODO-REPLACE: Add personal family acknowledgement.
[Author's personal acknowledgement to family — to be completed.]

% -------------------------------------------------------------------
\subsection*{Supervisor}
% TODO-REPLACE: Name supervisor(s) and describe their contribution.
[Supervisor name(s), institution, nature of guidance — to be completed.]

% -------------------------------------------------------------------
\subsection*{Trinity Hive Agents}
% TODO-REPLACE: Describe and properly attribute the autonomous agent
%               contributions per the university AI-use policy.
Beyond these public debts, the author thanks the maintainers of the
\textsf{tectonic}, \textsf{biblatex}, and \textsf{thmtools} projects
without which a Rust-only LaTeX pipeline would not have been possible
under R1 of the Trinity discipline.

% -------------------------------------------------------------------
\subsection*{Funding}
% TODO-REPLACE: Insert grant numbers and funding body acknowledgements.
[Funding body, grant reference number — to be completed.]

\bigskip
The errors that remain are the author's own. The ones that have been
caught and recorded are debts to those listed above.
