% ===================================================================
% Front matter — Preface
% ~2 pages, Springer LNM convention.
% ===================================================================

\chapter*{Preface}
\addcontentsline{toc}{chapter}{Preface}
\thispagestyle{plain}

This monograph began as a single algebraic observation: that the
positive root of \(x^{2} - x - 1 = 0\) satisfies
\(\varphi^{2} + \varphi^{-2} = 3\). The identity is elementary, and
no part of this work claims otherwise. What the work does claim is
that this elementary identity, taken seriously, has unusually long
reach: into the closure properties of the Lucas ring
\(\mathcal{L} = \mathbb{Z}[\varphi]\); into the \(\mathrm{E}_{8}\)
spectrum measured by Coldea and collaborators in CoBr\(_{2}\); into
the diffraction peaks of icosahedral quasicrystals; and into the
search space of a small neural architecture trained under
Coq-bounded constants.

The dissertation is organised so that the reader can take any of three
slices first. The \emph{theoretical} slice (Chapters~1--7 and 28--30)
develops the algebra. The \emph{empirical} slice (Chapters~8, 9,
20--26) records the falsifiable claims. The \emph{geometric} slice
(Chapters~11--19) gives the sacred-geometric reading of \(\varphi\)
that motivated the project before its formal-methods phase began.
A reader pressed for time may skip to Chapter~27 (\emph{Trinity
Identity}) and Chapter~24 (\emph{IGLA Architecture}); a reader
arriving from the formal-methods community may skip directly to
Appendix~F (\emph{Coq Citation Map}).

\subsection*{Editorial discipline}

Three editorial commitments structure every chapter of this work.

First, \emph{honest \texttt{Admitted}}. Where a Coq theorem in
\href{https://github.com/gHashTag/trinity-clara/tree/main/proofs/igla}{\texttt{trinity-clara/proofs/igla/}}
is mechanized but not yet closed, the corresponding monograph passage
carries an \texttt{\textbackslash admittedbox\{\ldots\}} marker that
the build system rejects to remove. There is no \emph{de facto}
upgrading of \texttt{Admitted} to \texttt{Proven} anywhere in the
text or in
\href{https://github.com/gHashTag/trios/blob/main/assertions/igla_assertions.json}{\texttt{assertions/igla\_assertions.json}}.

Second, \emph{traceability of constants} (rule L-R14). Every numeric
constant that appears in the monograph traces to a primary source
(CODATA, PDG, or a peer-reviewed Q1/Q2 journal), to a Coq theorem in
the \texttt{.v} files, and to a runtime callsite in
\texttt{crates/trios-*/src/}. Untraced constants cause the audit
binary to exit non-zero.

Third, \emph{falsifiability}. Every empirical chapter ends with an
explicit falsification criterion (Popper, 1959): a statement of the
observation that would refute the chapter's main claim, together with
the corroboration record to date. These records are aggregated in
Appendix~B.

\subsection*{Reproducibility}

The artefact pack accompanying the monograph targets the three ACM
Artifact Evaluation badges (\emph{Functional}, \emph{Reusable},
\emph{Available}). All experimental tables in Chapters~24--26 are
reproducible via
\verb|cargo run -p trios-phd reproduce --chapter <NN>| on a stated
hardware profile and seed list \(\{17, 42, 1729\}\). The data and
artefacts are deposited on Zenodo with a citable DOI; see Appendix~G.

\subsection*{What this work does not claim}

Two negative claims belong in the preface. This work does not claim
that \(\varphi\) is the \emph{only} structural principle in the
phenomena it touches. It claims, more narrowly, that \(\varphi\) is
\emph{a} sufficient principle, in a small set of well-defined
contexts, with mechanized proof and falsifiable empirical anchors.
The work also does not claim that the Coq mechanization is complete:
the \texttt{Admitted} list is non-empty, and the monograph names
each \texttt{Admitted} theorem honestly in Appendix~F.

\bigskip
\noindent\textit{Dmitrii Vasilev}\\
\textit{Trinity Research Programme, 2026}
