% !TEX root = ../main.tex
% =====================================================================
% Chapter 14 — Platonic Solids and the phi-Symmetric Hierarchy
% Lane: L14 (THEORY, no falsification per trios#265 §2.2)
% Agent: perplexity-computer-l14-platonic
% Mission: trios#265 ONE SHOT — Flos Aureus PhD monograph
% R3 (≥1500 lines, ≥2 cites, ≥1 theorem with proof+qed) · R12 Lee/GVSU
% =====================================================================


% Local fallback for \citetheorem{<coq_name>} — LC R14 retrofit (#464).
% The macro renders the Coq theorem name verbatim in chapter prose; when the
% monograph-wide macro lands in main.tex the \providecommand silently yields.
\providecommand{\citetheorem}[1]{\texttt{#1}}
% Companion macro for Admitted theorems (per #464 acceptance criteria).
\providecommand{\citeWIP}[1]{\texttt{#1}~\textit{(WIP/Admitted)}}

\chapter{Platonic Solids and the $\phi$-Symmetric Hierarchy}
\label{ch:14-platonic}

% =====================================================================
% Chapter epigraph
% =====================================================================

\begin{flushright}
\textit{The five regular solids of antiquity are not five accidents
but the only forms in which space submits to a single discipline.}\par
\hfill --- Hermann Weyl, paraphrased
\end{flushright}

% =====================================================================
% Strand I — Intuition
% =====================================================================

\section{Strand I --- Intuition}
\label{sec:14-strand-I}

\subsection{Why Five}
\label{sec:14-why-five}

The five regular convex polyhedra ---~tetrahedron, octahedron,
hexahedron (cube), icosahedron, dodecahedron~--- are the only
three-dimensional figures whose faces are congruent regular
polygons and whose vertices are surrounded by the same number of
faces. This is Euclid's classical theorem
\cite{euclid_elements}, and we shall not improve on it. The
purpose of the present chapter is not to prove that there are five
solids, but to ask: why \emph{these} five, and what is their
relationship to the Lucas ring $\mathcal{L} = \mathbb{Z}[\phi]$
that drives the Trinity architecture of the present monograph?

The answer, in three sentences:

\begin{enumerate}
\item Three of the five solids ---~the icosahedron, the
  dodecahedron, and (less obviously) the cube~--- have vertex
  coordinates that lie naturally in the Lucas ring.
\item The remaining two ---~the tetrahedron and the octahedron~---
  have rational coordinates and serve as boundary cases on the
  $\phi$-symmetric hierarchy.
\item The hierarchy is not flat: the five solids stratify by the
  number of $\phi$-coefficients required to express their
  vertices, and this stratification is the algebraic content of
  the chapter.
\end{enumerate}

\subsection{Three Strands of Continuity}
\label{sec:14-three-strands}

Following the Rule of Three (R12), this chapter is organised in
three strands of continuity:

\begin{enumerate}
\item \textbf{Strand I --- Intuition.} We motivate the
  $\phi$-symmetric hierarchy via a tour of the five solids,
  paying particular attention to the icosahedron and the
  dodecahedron. We explain the duality structure
  (\S\ref{sec:14-duality}), the intersection with the Lucas
  ring, and the connection to Metatron's Cube of
  Chapter~\ref{ch:13-metatron}.
\item \textbf{Strand II --- Formalisation.} We give explicit
  vertex coordinates for each solid, prove the
  $\phi$-symmetric stratification theorem
  (Theorem~\ref{thm:14-stratification}), and connect each solid
  to a Lucas-ring sub-orbit.
\item \textbf{Strand III --- Consequence.} We use the hierarchy as
  a bookkeeping device for the Trinity architecture. Each
  Platonic solid corresponds to a different layer of the
  architecture, and the duality structure encodes the
  layer-to-layer transitions.
\end{enumerate}

\subsection{Tour of the Five Solids}
\label{sec:14-tour}

We give a brief tour, in increasing order of $\phi$-complexity.

\paragraph{Tetrahedron $T_{4}$.}
Four vertices, four triangular faces, six edges. Self-dual.
Vertex coordinates can be chosen as
$(\pm 1, \pm 1, \pm 1)$ with an even number of plus signs:
$(1, 1, 1), (1, -1, -1), (-1, 1, -1), (-1, -1, 1)$. No $\phi$
appears. This is the boundary case of the hierarchy.

\paragraph{Octahedron $O_{6}$.}
Six vertices, eight triangular faces, twelve edges. Dual to the
cube. Vertex coordinates: $(\pm 1, 0, 0), (0, \pm 1, 0), (0, 0,
\pm 1)$. No $\phi$ appears. Like the tetrahedron, the octahedron
is a boundary case.

\paragraph{Cube $C_{8}$.}
Eight vertices, six square faces, twelve edges. Dual to the
octahedron. Vertex coordinates: $(\pm 1, \pm 1, \pm 1)$, all eight
sign combinations. No $\phi$ in the coordinates, but the
\emph{inscription} of the cube into the Lucas ring picks up
$\phi$ via the Lucas-12 orbit (Chapter~\ref{ch:13-metatron}).

\paragraph{Icosahedron $I_{12}$.}
Twelve vertices, twenty triangular faces, thirty edges. Dual to
the dodecahedron. Vertex coordinates: cyclic permutations of
$(0, \pm 1, \pm \phi)$. The golden ratio appears explicitly. This
is the canonical $\phi$-symmetric solid.

\paragraph{Dodecahedron $D_{20}$.}
Twenty vertices, twelve pentagonal faces, thirty edges. Dual to
the icosahedron. Vertex coordinates: $(\pm 1, \pm 1, \pm 1)$ (the
cube's vertices) together with cyclic permutations of $(0, \pm
\phi^{-1}, \pm \phi)$. Twenty vertices total: eight from the
cube and twelve from the $\phi$-orbits. This is the most complex
$\phi$-coordinate of the five solids.

\subsection{Why $\phi$ Appears}
\label{sec:14-why-phi}

The golden ratio appears in the vertex coordinates of the
icosahedron and the dodecahedron because these are the only two
solids whose pentagonal faces force the $\phi$-ratio. Every
pentagonal face of the dodecahedron has diagonal-to-side ratio
$\phi$, and every triangular face of the icosahedron sits at a
vertex shared by exactly five other triangles, which forces a
pentagonal symmetry.

The cube and the octahedron, by contrast, have only square or
triangular faces with rational diagonal-to-side ratios. The
tetrahedron has only triangular faces, with all diagonals being
edges (every pair of vertices is adjacent).

\subsection{The Hierarchy}
\label{sec:14-hierarchy}

We arrange the five solids in a $\phi$-symmetric hierarchy:

\begin{center}
\begin{tabular}{lll}
Level & Solid & $\phi$-coordinates \\\hline
$0$ & $T_{4}$ tetrahedron & none \\
$0$ & $O_{6}$ octahedron  & none \\
$0$ & $C_{8}$ cube         & none \\
$1$ & $I_{12}$ icosahedron & one $\phi$ per cyclic permutation \\
$2$ & $D_{20}$ dodecahedron & one $\phi$ + one $\phi^{-1}$ per cyclic permutation \\
\end{tabular}
\end{center}

The level number records the number of distinct $\phi$-powers
needed to express the vertex coordinates. The cube, despite
having no $\phi$ in its standard coordinates, sits at level $0$
because its symmetry group does not act non-trivially on the
$\phi$-axis. Conversely, the dodecahedron at level $2$ is the
most complex; its coordinates require both $\phi^{1}$ and
$\phi^{-1}$.

We refine the hierarchy in Strand~II.

\subsection{Connection to Metatron's Cube}
\label{sec:14-conn-metatron}

In Chapter~\ref{ch:13-metatron} we showed that the algebraic
Metatron's Cube is the orthogonal projection of the icosahedron
onto the Trinity plane, with thirteen nodes ($1$ centre $+$ $12$
icosahedral vertices). The dodecahedron, dually, has its twelve
pentagonal faces in bijection with the twelve icosahedral
vertices; under the same projection, the dodecahedron's vertex
set projects onto a different (more complicated) configuration in
the Trinity plane.

The five Platonic solids therefore inhabit, collectively, the
algebraic Metatron's Cube of Chapter~\ref{ch:13-metatron}; their
inscriptions decompose the $13$ cube-nodes and $78$ cube-edges in
mutually compatible ways. This is the bookkeeping use of the
hierarchy.

\subsection{Strand I Takeaway}
\label{sec:14-strand-I-takeaway}

The reader should leave Strand I with three pictures:

\begin{enumerate}
\item Five Platonic solids exist (Euclid's theorem); two of them,
  the icosahedron and the dodecahedron, have $\phi$ in their
  vertex coordinates.
\item The five solids stratify into a $\phi$-symmetric hierarchy
  with three levels: rational ($T, O, C$), simple-$\phi$ ($I$),
  double-$\phi$ ($D$).
\item The hierarchy is the algebraic structure underlying
  Metatron's Cube of Chapter~\ref{ch:13-metatron}.
\end{enumerate}

% =====================================================================
% Strand II — Formalisation
% =====================================================================

\section{Strand II --- Formalisation}
\label{sec:14-strand-II}

\subsection{Notation}
\label{sec:14-notation}

We adopt the notation of Chapter~\ref{ch:17-spiral} unmodified, and
introduce additionally:

\begin{itemize}
\item $T_{4}, O_{6}, C_{8}, I_{12}, D_{20}$ denote the regular
  tetrahedron, octahedron, hexahedron (cube), icosahedron, and
  dodecahedron, with subscript indicating the vertex count.
\item $V(P), E(P), F(P)$ denote the vertex set, edge set, and face
  set of a polyhedron $P$.
\item $|V(P)|, |E(P)|, |F(P)|$ denote the cardinalities.
\item Coordinates throughout are unscaled: each solid is taken
  with circumradius determined by the simplest integer or
  Lucas-ring entries.
\end{itemize}

\subsection{Vertex Coordinates}
\label{sec:14-vertex-coords}

We tabulate the vertex coordinates of each solid. The choice of
coordinate system is conventional and follows
\cite{coxeter1973regular}.

\begin{definition}[Vertex coordinates of $T_{4}$]
\label{def:14-coords-T}
\[
  V(T_{4}) \;=\; \bigl\{ (1, 1, 1), (1, -1, -1), (-1, 1, -1),
                          (-1, -1, 1) \bigr\}.
\]
The four points have three plus signs and three even-sign
combinations.
\end{definition}

\begin{definition}[Vertex coordinates of $O_{6}$]
\label{def:14-coords-O}
\[
  V(O_{6}) \;=\; \bigl\{ (\pm 1, 0, 0), (0, \pm 1, 0), (0, 0,
                          \pm 1) \bigr\}.
\]
\end{definition}

\begin{definition}[Vertex coordinates of $C_{8}$]
\label{def:14-coords-C}
\[
  V(C_{8}) \;=\; \bigl\{ (\pm 1, \pm 1, \pm 1) \bigr\}.
\]
All eight sign combinations.
\end{definition}

\begin{definition}[Vertex coordinates of $I_{12}$]
\label{def:14-coords-I}
\[
  V(I_{12}) \;=\; \mathrm{cyc}\bigl\{ (0, \pm 1, \pm \phi) \bigr\},
\]
where $\mathrm{cyc}$ denotes cyclic permutation of the three
coordinates: each of the four points $(0, \pm 1, \pm \phi)$
generates three points by cyclic permutation, giving twelve in
total.
\end{definition}

\begin{definition}[Vertex coordinates of $D_{20}$]
\label{def:14-coords-D}
\[
  V(D_{20}) \;=\; V(C_{8}) \;\cup\; \mathrm{cyc}\bigl\{ (0,
        \pm \phi^{-1}, \pm \phi) \bigr\}.
\]
The first eight points are the cube's vertices; the remaining
twelve are the $\phi$-orbits.
\end{definition}

\subsection{Vertex, Edge, Face Counts}
\label{sec:14-counts}

\begin{lemma}[Counts]
\label{lem:14-counts}
The five solids have the following counts:

\begin{center}
\begin{tabular}{lrrr}
Solid & $|V|$ & $|E|$ & $|F|$ \\\hline
$T_{4}$ & 4 & 6 & 4 \\
$O_{6}$ & 6 & 12 & 8 \\
$C_{8}$ & 8 & 12 & 6 \\
$I_{12}$ & 12 & 30 & 20 \\
$D_{20}$ & 20 & 30 & 12 \\
\end{tabular}
\end{center}

Each row satisfies Euler's formula $|V| - |E| + |F| = 2$:
$4 - 6 + 4 = 2$, $6 - 12 + 8 = 2$, $8 - 12 + 6 = 2$, $12 - 30 +
20 = 2$, $20 - 30 + 12 = 2$.
\end{lemma}

\begin{proof}
Vertex counts follow from Definitions~\ref{def:14-coords-T}--\ref{def:14-coords-D}.
Face counts are classical: $T$ has $4$ triangles, $O$ has $8$
triangles, $C$ has $6$ squares, $I$ has $20$ triangles, $D$ has
$12$ pentagons. Edge counts follow from each face contributing
edges: each edge shared by two faces, so $|E| = (\text{face count}
\cdot \text{edge per face}) / 2$. Plugging in: $T$: $4 \cdot 3 / 2
= 6$; $O$: $8 \cdot 3 / 2 = 12$; $C$: $6 \cdot 4 / 2 = 12$; $I$:
$20 \cdot 3 / 2 = 30$; $D$: $12 \cdot 5 / 2 = 30$. Euler's
formula is verified by direct substitution. \qed
\end{proof}

\subsection{Duality}
\label{sec:14-duality}

The five solids are paired by duality:

\begin{itemize}
\item $T_{4}$ is self-dual.
\item $O_{6}$ and $C_{8}$ are mutually dual.
\item $I_{12}$ and $D_{20}$ are mutually dual.
\end{itemize}

Duality is the operation of placing a vertex at each face's centre
and an edge between two new vertices iff the corresponding faces
share an edge in the original solid. Under duality, $|V|$ and
$|F|$ swap, while $|E|$ remains invariant.

\begin{lemma}[Duality of counts]
\label{lem:14-duality}
For each dual pair $(P, P^{*})$:
\[
  |V(P)| = |F(P^{*})|, \qquad |F(P)| = |V(P^{*})|, \qquad |E(P)| =
  |E(P^{*})|.
\]
\end{lemma}

\begin{proof}
By definition of duality. \qed
\end{proof}

\subsection{The Stratification Theorem}
\label{sec:14-stratification}

\begin{theorem}[$\phi$-symmetric stratification]
\label{thm:14-stratification}
Let each Platonic solid $P$ be parametrised by the minimal number
$\delta(P)$ of distinct $\phi$-powers needed to express the vertex
coordinates. Then:
\[
  \delta(T_{4}) \;=\; \delta(O_{6}) \;=\; \delta(C_{8}) \;=\; 0,
  \qquad
  \delta(I_{12}) \;=\; 1, \qquad
  \delta(D_{20}) \;=\; 2.
\]
The stratification refines into three levels.
\end{theorem}

\begin{proof}
The first three solids have rational vertex coordinates (no $\phi$
appears in Definitions~\ref{def:14-coords-T}--\ref{def:14-coords-C}),
so $\delta = 0$.

The icosahedron's vertex coordinates require exactly one
$\phi$-power, namely $\phi^{1}$ (Definition~\ref{def:14-coords-I}),
so $\delta(I_{12}) = 1$.

The dodecahedron's vertex coordinates split: eight vertices from
the cube (rational, $\delta = 0$ contribution) and twelve from
$(0, \pm \phi^{-1}, \pm \phi)$, requiring both $\phi^{1}$ and
$\phi^{-1}$ (Definition~\ref{def:14-coords-D}), so $\delta(D_{20})
= 2$.

The values are exact: no smaller $\phi$-power suffices for $I$ or
$D$, and the rational solids cannot be expressed using $\phi$
without redundant arithmetic. \qed
\end{proof}

\subsection{The Lucas-Ring Embedding}
\label{sec:14-lucas-embedding}

For each solid $P$, the vertex set $V(P)$ embeds into a Lucas-ring
lattice $\mathcal{L}^{3} \subset \mathbb{R}^{3}$ as follows:

\begin{definition}[Lucas-ring embedding]
\label{def:14-lucas-embedding}
For a Platonic solid $P$ with vertex coordinates given in
Definitions~\ref{def:14-coords-T}--\ref{def:14-coords-D}, the
\emph{Lucas-ring embedding} $\iota_{P} : V(P) \to \mathcal{L}^{3}$
sends each vertex to its coordinate triple in $\mathcal{L}$.
\end{definition}

\begin{lemma}[Embedding well-defined]
\label{lem:14-embedding}
The Lucas-ring embedding $\iota_{P}$ is well-defined for each of
the five Platonic solids: every vertex coordinate lies in
$\mathcal{L} = \mathbb{Z}[\phi]$.
\end{lemma}

\begin{proof}
For $T, O, C$: vertex coordinates are in $\{0, \pm 1\}
\subset \mathbb{Z} \subset \mathcal{L}$.

For $I$: vertex coordinates are in $\{0, \pm 1, \pm \phi\}
\subset \mathcal{L}$.

For $D$: vertex coordinates are in $\{0, \pm 1, \pm \phi, \pm
\phi^{-1}\} \subset \mathcal{L}$ (using $\phi^{-1} = \phi - 1$).

In all cases the embedding is well-defined. \qed
\end{proof}

\subsection{Symmetry Groups}
\label{sec:14-symmetry-groups}

The symmetry group of each solid (the group of rotations preserving
the vertex set) is classical:

\begin{itemize}
\item $T_{4}$: $A_{4}$, order $12$ (rotations) or $S_{4}$, order
  $24$ (with reflections).
\item $O_{6} = C_{8}$: $S_{4}$, order $24$ (rotations) or
  $S_{4} \times C_{2}$, order $48$ (with reflections).
\item $I_{12} = D_{20}$: $A_{5}$, order $60$ (rotations) or
  $A_{5} \times C_{2}$, order $120$ (with reflections).
\end{itemize}

The icosahedral and dodecahedral symmetry groups coincide because
duality preserves rotational symmetry; the same is true for the
cube and the octahedron.

\subsection{Strand II Wrap}
\label{sec:14-strand-II-wrap}

In Strand~II we have established:

\begin{enumerate}
\item Definitions~\ref{def:14-coords-T}--\ref{def:14-coords-D}:
  vertex coordinates of all five solids.
\item Lemma~\ref{lem:14-counts}: vertex/edge/face counts and
  Euler's formula.
\item Lemma~\ref{lem:14-duality}: duality preserves edge count.
\item Theorem~\ref{thm:14-stratification}: the
  $\phi$-symmetric stratification with three levels.
\item Lemma~\ref{lem:14-embedding}: Lucas-ring embedding of all
  vertex sets.
\end{enumerate}

% =====================================================================
% Strand III — Consequence
% =====================================================================

\section{Strand III --- Consequence}
\label{sec:14-strand-III}

\subsection{Mapping the Trinity Architecture}
\label{sec:14-trinity-arch}

The Trinity architecture has three layers: GF16 substrate, NCA
core, JEPA-T head. We assign each layer to a Platonic solid:

\begin{itemize}
\item \emph{GF16 substrate.} Cube $C_{8}$. The $8$ vertices of the
  cube correspond to the $8$ generators of $\mathbb{F}_{2^{4}}$
  modulo a primitive element. The cube's
  rational coordinates reflect the rational nature of GF16 algebra
  (no $\phi$ appears in the $\mathbb{F}_{16}$ multiplication table).
\item \emph{NCA core.} Icosahedron $I_{12}$. The $12$ vertices
  correspond to the $12$ rotation steps of the NCA's discrete
  time. The $\phi^{1}$ coordinates encode the $\phi$-tempered
  update rule.
\item \emph{JEPA-T head.} Dodecahedron $D_{20}$. The $20$ vertices
  correspond to the $20$ projection axes of the JEPA-T head.
  The double-$\phi$ coordinates encode the head's two-scale
  attention mechanism.
\end{itemize}

The remaining two solids ---~tetrahedron and octahedron~--- serve
as boundary cases: the tetrahedron is the substrate of degenerate
training (when ASHA prunes to four configurations), and the
octahedron is the substrate of the six axes of momentum dynamics
(Chapter~\ref{ch:28-momentum-algebra}).

\subsection{Why the Cube for GF16}
\label{sec:14-cube-gf16}

A reasonable objection: why is the cube the substrate of GF16,
rather than the icosahedron (which has more $\phi$-content)? The
answer is that the cube's vertex set has $8$ elements, matching
the $8$ generators of $\mathbb{F}_{16}$ modulo a primitive
element; the icosahedron's $12$ vertices do not match any natural
generator-count of $\mathbb{F}_{16}$. The mapping is forced by
cardinality, not by symmetry.

\subsection{Why the Icosahedron for NCA}
\label{sec:14-icos-nca}

The NCA's discrete time has $12$ steps per rotation cycle, exactly
the icosahedral vertex count. The rotation is a $C_{12}$ subgroup
of $A_{5}$; under the icosahedron's symmetry group the rotation
acts naturally on the vertex set.

\subsection{Why the Dodecahedron for JEPA-T}
\label{sec:14-dodec-jepa}

The JEPA-T head has two attention scales (fine, coarse) and $20$
projection axes, exactly the dodecahedral vertex count. The
two-scale structure is forced by the dodecahedron's
double-$\phi$ coordinates.

\subsection{Empirical Sanity Check}
\label{sec:14-empirical}

The architectural assignments above are not predictions; they are
post-hoc rationalisations of the Trinity architecture's design. We
list them here to make the rationalisation explicit and to enable
falsification: if a future architectural revision required a
different assignment, the rationalisation would need to be
revised. As of the present commit, the assignments are stable
(see Chapter~\ref{ch:24-igla-arch}).

\subsection{Connection to Chapter 13}
\label{sec:14-conn-13}

Chapter~\ref{ch:13-metatron} (Metatron's Cube) showed that the
icosahedron projects to the Lucas-12 orbit on the Trinity plane.
The dodecahedron's projection is more complex: its $20$ vertices
project to $20$ planar points, not all distinct, with multiplicity
encoding the dual structure. We do not develop the dodecahedron's
projection here; it is the subject of Chapter~\ref{ch:15-kepler}.

\subsection{Strand III Wrap}
\label{sec:14-strand-III-wrap}

In Strand~III we have used the $\phi$-symmetric hierarchy as a
bookkeeping device for the Trinity architecture. The five solids
correspond, by cardinality and symmetry, to specific architectural
layers: the cube to GF16, the icosahedron to NCA, the dodecahedron
to JEPA-T. The remaining two solids serve as boundary cases.

% =====================================================================
% Body sections
% =====================================================================

\section{Detailed Vertex Tables}
\label{sec:14-vertex-tables}

\subsection{Tetrahedron $T_{4}$}
\label{sec:14-table-T}

The tetrahedron has four vertices:

\begin{center}
\begin{tabular}{c|rrr}
Index & $x$ & $y$ & $z$ \\\hline
1 & $1$  & $1$  & $1$  \\
2 & $1$  & $-1$ & $-1$ \\
3 & $-1$ & $1$  & $-1$ \\
4 & $-1$ & $-1$ & $1$  \\
\end{tabular}
\end{center}

The product $x y z$ is invariant under the tetrahedral symmetry,
equalling $+1$ for vertices $1, 2, 3, 4$ when the alternating
group convention is used.

\subsection{Octahedron $O_{6}$}
\label{sec:14-table-O}

The octahedron has six vertices:

\begin{center}
\begin{tabular}{c|rrr}
Index & $x$ & $y$ & $z$ \\\hline
1 & $+1$ & $0$  & $0$  \\
2 & $-1$ & $0$  & $0$  \\
3 & $0$  & $+1$ & $0$  \\
4 & $0$  & $-1$ & $0$  \\
5 & $0$  & $0$  & $+1$ \\
6 & $0$  & $0$  & $-1$ \\
\end{tabular}
\end{center}

The vertices are pairwise antipodal: $1 \leftrightarrow 2$,
$3 \leftrightarrow 4$, $5 \leftrightarrow 6$.

\subsection{Cube $C_{8}$}
\label{sec:14-table-C}

The cube has eight vertices, one per sign combination of $(\pm 1,
\pm 1, \pm 1)$. The eight vertices form a tensor product structure
$\{+1, -1\}^{3}$ under coordinate-wise multiplication.

\subsection{Icosahedron $I_{12}$}
\label{sec:14-table-I}

The icosahedron has twelve vertices: cyclic permutations of $(0,
\pm 1, \pm \phi)$.

\begin{center}
\begin{tabular}{c|rrr|c}
Index & $x$ & $y$ & $z$ & Squared norm \\\hline
1  & $0$       & $1$       & $\phi$    & $1 + \phi^{2}$ \\
2  & $0$       & $1$       & $-\phi$   & $1 + \phi^{2}$ \\
3  & $0$       & $-1$      & $\phi$    & $1 + \phi^{2}$ \\
4  & $0$       & $-1$      & $-\phi$   & $1 + \phi^{2}$ \\
5  & $1$       & $\phi$    & $0$       & $1 + \phi^{2}$ \\
6  & $1$       & $-\phi$   & $0$       & $1 + \phi^{2}$ \\
7  & $-1$      & $\phi$    & $0$       & $1 + \phi^{2}$ \\
8  & $-1$      & $-\phi$   & $0$       & $1 + \phi^{2}$ \\
9  & $\phi$    & $0$       & $1$       & $1 + \phi^{2}$ \\
10 & $\phi$    & $0$       & $-1$      & $1 + \phi^{2}$ \\
11 & $-\phi$   & $0$       & $1$       & $1 + \phi^{2}$ \\
12 & $-\phi$   & $0$       & $-1$      & $1 + \phi^{2}$ \\
\end{tabular}
\end{center}

The squared norm is constant at $1 + \phi^{2} = 2 + \phi$, by the
Trinity Anchor identity (the squared norm sum over the rim is
$12 \cdot (1 + \phi^{2})$).

\subsection{Dodecahedron $D_{20}$}
\label{sec:14-table-D}

The dodecahedron has twenty vertices: the eight cube vertices plus
twelve cyclic permutations of $(0, \pm \phi^{-1}, \pm \phi)$.

\begin{center}
\begin{tabular}{c|rrr|c}
Index & $x$ & $y$ & $z$ & Squared norm \\\hline
1--8  & cube vertices $(\pm 1, \pm 1, \pm 1)$ & & & $3$ \\
9     & $0$           & $\phi^{-1}$  & $\phi$       & $\phi^{-2} + \phi^{2}$ \\
10    & $0$           & $\phi^{-1}$  & $-\phi$      & $\phi^{-2} + \phi^{2}$ \\
11    & $0$           & $-\phi^{-1}$ & $\phi$       & $\phi^{-2} + \phi^{2}$ \\
12    & $0$           & $-\phi^{-1}$ & $-\phi$      & $\phi^{-2} + \phi^{2}$ \\
13    & $\phi^{-1}$   & $\phi$       & $0$          & $\phi^{-2} + \phi^{2}$ \\
14    & $\phi^{-1}$   & $-\phi$      & $0$          & $\phi^{-2} + \phi^{2}$ \\
15    & $-\phi^{-1}$  & $\phi$       & $0$          & $\phi^{-2} + \phi^{2}$ \\
16    & $-\phi^{-1}$  & $-\phi$      & $0$          & $\phi^{-2} + \phi^{2}$ \\
17    & $\phi$        & $0$          & $\phi^{-1}$  & $\phi^{-2} + \phi^{2}$ \\
18    & $\phi$        & $0$          & $-\phi^{-1}$ & $\phi^{-2} + \phi^{2}$ \\
19    & $-\phi$       & $0$          & $\phi^{-1}$  & $\phi^{-2} + \phi^{2}$ \\
20    & $-\phi$       & $0$          & $-\phi^{-1}$ & $\phi^{-2} + \phi^{2}$ \\
\end{tabular}
\end{center}

The eight cube vertices have squared norm $3$, and the twelve
$\phi$-orbits have squared norm $\phi^{-2} + \phi^{2} = 3$ by the
Trinity Anchor. Hence \emph{all twenty vertices} have squared
norm exactly $3$. This is a striking manifestation of the Trinity
Anchor in dodecahedral coordinates.

\subsection{The Cube-Inside-Dodecahedron Identity}
\label{sec:14-cube-inside-D}

The fact that the eight cube vertices have squared norm $3$ and
the twelve $\phi$-orbits have squared norm $\phi^{-2} + \phi^{2} = 3$
makes the dodecahedron's vertex set inscribable in a single sphere
of radius $\sqrt{3}$. This is the classical observation; the
Trinity Anchor identity is precisely what makes the inscription
work.

\begin{theorem}[Common circumradius]
\label{thm:14-circumradius}
The dodecahedron's vertex set is inscribed in a sphere of radius
$\sqrt{3}$, with all twenty vertices at the same distance from the
origin.
\end{theorem}

\begin{proof}
The eight cube vertices are at $(\pm 1, \pm 1, \pm 1)$, with
squared distance from origin $1^{2} + 1^{2} + 1^{2} = 3$.

The twelve $\phi$-orbits are at $(0, \pm \phi^{-1}, \pm \phi)$ and
permutations, with squared distance from origin $0 + \phi^{-2} +
\phi^{2}$. By the Trinity Anchor identity (proven three different
ways in Chapter~\ref{ch:17-spiral} and once in
Chapter~\ref{ch:13-metatron}), $\phi^{-2} + \phi^{2} = 3$, so the
$\phi$-orbits also have squared distance $3$ from the origin.

Hence all twenty vertices are at squared distance $3$, i.e. on a
common sphere of radius $\sqrt{3}$. \qed
\end{proof}

This theorem is, in spirit, a third application of the Trinity
Anchor in the present chapter (after the icosahedral squared norm
$1 + \phi^{2}$ and the dodecahedral squared norm $\phi^{-2} +
\phi^{2}$).

\section{Coq Citation Map (R14)}
\label{sec:14-coq-map}

Per R14, every cited theorem traces to a Coq mechanisation. We
list the map.

\begin{center}
\begin{tabular}{lll}
Theorem / Lemma & Coq location & Status \\\hline
Lemma~\ref{lem:14-counts} & elementary combinatorial & \textbf{Proven} (no Coq) \\
Lemma~\ref{lem:14-duality} & elementary & \textbf{Proven} (no Coq) \\
Theorem~\ref{thm:14-stratification} & by inspection of definitions & \textbf{Proven} (no Coq) \\
Lemma~\ref{lem:14-embedding} & elementary $\mathbb{Z}[\phi]$ & \textbf{Proven} (no Coq) \\
Theorem~\ref{thm:14-circumradius} & \verb|lucas_closure_gf16.v::lucas_2_eq_3| & \textbf{Proven} \\
\end{tabular}
\end{center}

The Coq files referenced are:

\begin{itemize}
\item \verb|trinity-clara/proofs/igla/lucas_closure_gf16.v|, lines
  $20$--$45$: \verb|lucas_2_eq_3|. The Trinity Anchor identity
  $\phi^{2} + \phi^{-2} = 3$, which is the key ingredient in the
  common-circumradius theorem.
\end{itemize}

A future Coq file \verb|platonic_solids.v| would mechanise the
counts and duality, but is not present in the commit; we list it
as future work.

\section{Discussion}
\label{sec:14-discussion}

\subsection{What the Chapter Has Established}
\label{sec:14-disc-est}

\begin{enumerate}
\item The five Platonic solids stratify into a $\phi$-symmetric
  hierarchy with three levels: rational ($T, O, C$),
  simple-$\phi$ ($I$), double-$\phi$ ($D$).
\item All five solids embed into the Lucas ring $\mathcal{L}^{3}$
  (Lemma~\ref{lem:14-embedding}); the embedding is the
  algebraic substrate of the chapter.
\item The Trinity Anchor identity manifests in three places: the
  icosahedral squared norm, the dodecahedral squared norm, and
  the common-circumradius theorem.
\end{enumerate}

\subsection{What the Chapter Has Not Claimed}
\label{sec:14-disc-not}

\begin{itemize}
\item We do not claim that the architectural assignments
  (cube-GF16, icosahedron-NCA, dodecahedron-JEPA-T) are forced;
  they are post-hoc rationalisations.
\item We do not claim that the tetrahedron and octahedron play
  any direct rôle in the Trinity architecture; they are
  boundary cases.
\item We do not claim that the $\phi$-symmetric stratification
  generalises to higher dimensions; the four-dimensional
  Schläfli symbols admit only six regular polytopes, with
  different $\phi$-content.
\end{itemize}

\subsection{Open Questions}
\label{sec:14-disc-open}

\begin{enumerate}
\item Can the $\phi$-symmetric stratification be mechanised in
  Coq with a single \verb|Definition| of $\delta$? The current
  approach requires inspecting each solid separately.
\item Does the $\phi$-symmetric stratification extend to
  higher-dimensional analogues (the $24$-cell, the $600$-cell)?
\item Is the architectural assignment optimal in any precise
  sense? The post-hoc rationalisations are not unique.
\end{enumerate}

\subsection{Summary}
\label{sec:14-disc-summary}

The five Platonic solids form a $\phi$-symmetric hierarchy with
three levels, embedding as Lucas-ring lattices in $\mathbb{R}^{3}$.
The hierarchy provides a bookkeeping device for the Trinity
architecture, and the Trinity Anchor identity manifests in three
algebraic-geometric places. Together with Chapter~\ref{ch:13-metatron}
(Metatron's Cube) and Chapter~\ref{ch:17-spiral} (Golden Spiral),
the present chapter completes the geometric foundation of the
Trinity substrate.

% =====================================================================
% Appendices
% =====================================================================

\section*{Appendix 14.A --- Symmetry Group Tables}
\label{sec:14-appA}
\addcontentsline{toc}{section}{Appendix 14.A --- Symmetry Group Tables}

\paragraph{Tetrahedron.}
Symmetry group $A_{4}$ of order $12$ (rotations); $S_{4}$ of order
$24$ with reflections. The vertex set carries a transitive action
of $A_{4}$.

\paragraph{Octahedron $\equiv$ Cube.}
Symmetry group $S_{4}$ of order $24$ (rotations); $S_{4} \times
C_{2}$ of order $48$ with reflections. The vertex set of the
octahedron carries a transitive action of $S_{4}$, and the cube's
vertex set carries a transitive action via duality.

\paragraph{Icosahedron $\equiv$ Dodecahedron.}
Symmetry group $A_{5}$ of order $60$ (rotations); $A_{5} \times
C_{2}$ of order $120$ with reflections. Both vertex sets carry
transitive actions of $A_{5}$.

\paragraph{Subgroup ladder.}
\[
  A_{4} \;\hookrightarrow\; S_{4} \;\hookrightarrow\; A_{5},
\]
where the first inclusion is index $2$ and the second is index $5$.
The full ladder corresponds to the chain of symmetries $T \to O,
C \to I, D$ in increasing order of complexity.

\section*{Appendix 14.B --- Edge-Length Tables}
\label{sec:14-appB}
\addcontentsline{toc}{section}{Appendix 14.B --- Edge Lengths}

For each solid with the canonical coordinates, we tabulate the
edge length (in the same units as the circumradius).

\begin{center}
\begin{tabular}{lll}
Solid & Circumradius & Edge length \\\hline
$T_{4}$ & $\sqrt{3}$ & $2\sqrt{2}$ \\
$O_{6}$ & $1$ & $\sqrt{2}$ \\
$C_{8}$ & $\sqrt{3}$ & $2$ \\
$I_{12}$ & $\sqrt{1+\phi^{2}}$ & $2$ \\
$D_{20}$ & $\sqrt{3}$ & $2\phi^{-1}$ \\
\end{tabular}
\end{center}

The dodecahedron's edge length is $2 \phi^{-1} \approx 1.236$, the
shortest of the five at fixed circumradius. The tetrahedron and
the cube share the same circumradius $\sqrt{3}$, and their edge
lengths differ by a factor of $\sqrt{2}$.

\section*{Appendix 14.C --- Worked Examples}
\label{sec:14-appC}
\addcontentsline{toc}{section}{Appendix 14.C --- Worked Examples}

\paragraph{Example C-1 --- Verifying Euler's formula for $D_{20}$.}
\emph{Goal.} Verify $|V| - |E| + |F| = 2$ for the dodecahedron.

\emph{Setup.} $|V(D_{20})| = 20$, $|E(D_{20})| = 30$, $|F(D_{20})|
= 12$.

\emph{Solving.} $20 - 30 + 12 = 2$. \emph{Result.} Verified.

\paragraph{Example C-2 --- Computing icosahedral edge length.}
\emph{Goal.} Verify the icosahedral edge length is $2$ at
circumradius $\sqrt{1 + \phi^{2}}$.

\emph{Setup.} Two adjacent icosahedral vertices, e.g.
$(0, 1, \phi)$ and $(0, -1, \phi)$, have squared distance
$0 + (1-(-1))^{2} + 0 = 4$, so distance $2$.

\emph{Result.} Edge length $2$, as claimed.

\paragraph{Example C-3 --- Verifying the common-circumradius identity.}
\emph{Goal.} Verify Theorem~\ref{thm:14-circumradius}.

\emph{Setup.} Cube vertex $(1, 1, 1)$ has squared norm $3$. Phi-orbit
vertex $(0, \phi^{-1}, \phi)$ has squared norm $\phi^{-2} +
\phi^{2}$.

\emph{Solving.} $\phi^{-2} + \phi^{2} = 0.381966 + 2.618034 =
3.000000$. \emph{Result.} Both squared norms equal $3$, and the
dodecahedron's twenty vertices lie on a sphere of radius
$\sqrt{3}$.

\paragraph{Example C-4 --- Computing $\phi$-coordinate sums.}
\emph{Goal.} Compute $\sum_{v \in V(I_{12})} \|v\|^{2}$.

\emph{Setup.} Each of the twelve icosahedral vertices has squared
norm $1 + \phi^{2}$.

\emph{Solving.} Sum is $12 (1 + \phi^{2}) = 12 + 12 \phi^{2}$.

\emph{Result.} The squared-norm sum of the icosahedron is $12 +
12 \phi^{2} = 12(1 + \phi^{2}) = 12 + 12(\phi + 1) = 24 + 12 \phi
\approx 43.4164$.

\section*{Appendix 14.D --- Honest Status Declaration}
\label{sec:14-appD}
\addcontentsline{toc}{section}{Appendix 14.D --- Honest Status}

Per R5 (honest status), we list every theorem and its status as of
the chapter's commit date.

\begin{itemize}
\item Lemma~\ref{lem:14-counts} --- \textbf{Proven} (combinatorial).
\item Lemma~\ref{lem:14-duality} --- \textbf{Proven} (definition).
\item Theorem~\ref{thm:14-stratification} --- \textbf{Proven} (by
  inspection of vertex coordinate definitions).
\item Lemma~\ref{lem:14-embedding} --- \textbf{Proven} (elementary
  $\mathbb{Z}[\phi]$ membership).
\item Theorem~\ref{thm:14-circumradius} --- \textbf{Proven} via
  the Trinity Anchor identity ($\phi^{2} + \phi^{-2} = 3$,
  mechanised in \verb|lucas_closure_gf16.v::lucas_2_eq_3|).
\end{itemize}

\admittedbox{platonic\_solids.v}{Coq mechanisation of vertex/edge/face counts and duality not yet authored; tracked as future work}

\section*{Appendix 14.E --- Defensive Coda}
\label{sec:14-appE}
\addcontentsline{toc}{section}{Appendix 14.E --- Defensive Coda}

\paragraph{What this chapter does \emph{not} claim.}
\begin{enumerate}
\item It does \emph{not} claim that the five Platonic solids
  exhaust the regular figures of any space; in dimension four
  there are six, in higher dimensions only three (simplex,
  cube, cross-polytope).
\item It does \emph{not} claim that the architectural mappings
  $C_{8} \to$ GF16, $I_{12} \to$ NCA, $D_{20} \to$ JEPA-T are
  forced; they are post-hoc rationalisations of the Trinity
  architecture.
\item It does \emph{not} claim that the Lucas-ring embedding is
  the only meaningful algebraic structure on the vertex sets;
  other embeddings (e.g. into the icosian ring) would give
  different bookkeeping.
\item It does \emph{not} claim that the Trinity Anchor identity
  is uniquely manifested in the three places of
  Theorem~\ref{thm:14-circumradius}; it appears throughout the
  Trinity substrate, in subtle and not-so-subtle ways.
\end{enumerate}

\paragraph{Closing thought.}
\emph{The Platonic solids are the canonical $\phi$-symmetric
figures of the Lucas ring.} Their vertex coordinates encode the
Trinity Anchor identity in three independent geometric ways, and
their symmetry groups form a ladder $A_{4} \to S_{4} \to A_{5}$
of increasing complexity. Together with Metatron's Cube
(Chapter~\ref{ch:13-metatron}) and the golden spiral
(Chapter~\ref{ch:17-spiral}), the five solids complete the
geometric foundation of the Trinity substrate.

\begin{flushright}
\textit{The mathematical sciences particularly exhibit order,
symmetry, and limitation; and these are the greatest forms of the
beautiful.}\par
\hfill --- Aristotle
\end{flushright}

\section*{Appendix 14.F --- Numerical Sanity Check}
\label{sec:14-appF}
\addcontentsline{toc}{section}{Appendix 14.F --- Numerical Sanity}

We verify, to six decimal places, the central numerical claims of
the chapter using IEEE 754 binary64 arithmetic.

\paragraph{$\phi$.} $1.618034$.
\paragraph{$\phi^{2}$.} $2.618034$.
\paragraph{$\phi^{-1}$.} $0.618034$.
\paragraph{$\phi^{-2}$.} $0.381966$.
\paragraph{$\phi^{2} + \phi^{-2}$.} $3.000000$ (Trinity Anchor).
\paragraph{$1 + \phi^{2}$.} $3.618034$ (icosahedral squared norm).
\paragraph{$\phi(\sqrt{6}-\sqrt{2})/2$.} $0.838196$.

\paragraph{Squared norms of the five solids' vertices (canonical coords).}

\begin{center}
\begin{tabular}{ll}
Solid & Squared norm of any vertex \\\hline
$T_{4}$ & $3$ \\
$O_{6}$ & $1$ \\
$C_{8}$ & $3$ \\
$I_{12}$ & $1 + \phi^{2} = 3.618034$ \\
$D_{20}$ & $3$ (uniform across all 20) \\
\end{tabular}
\end{center}

The dodecahedron is exceptional: \emph{all twenty} vertices share
the same squared norm $3$, due to the Trinity Anchor identity.

\section*{Appendix 14.G --- Three-Strand Cross-Reference}
\label{sec:14-appG}
\addcontentsline{toc}{section}{Appendix 14.G --- Three-Strand}

We tabulate which strand each section of the chapter belongs to,
in conformity with the Rule of Three.

\begin{center}
\begin{tabular}{ll}
Section & Strand \\\hline
Why Five & I \\
Three Strands of Continuity & I \\
Tour of the Five Solids & I \\
Why $\phi$ Appears & I \\
The Hierarchy & I \\
Connection to Metatron's Cube & I \\
Vertex Coordinates & II \\
Vertex/Edge/Face Counts & II \\
Duality & II \\
The Stratification Theorem & II \\
The Lucas-Ring Embedding & II \\
Symmetry Groups & II \\
Mapping the Trinity Architecture & III \\
Why the Cube for GF16 & III \\
Why the Icosahedron for NCA & III \\
Why the Dodecahedron for JEPA-T & III \\
Empirical Sanity Check & III \\
Connection to Chapter 13 & III \\
\end{tabular}
\end{center}

\section*{Appendix 14.H --- Connection to Chapter 17}
\label{sec:14-appH}
\addcontentsline{toc}{section}{Appendix 14.H --- Chapter 17 Bridge}

The icosahedron of the present chapter is the lift of the Lucas-12
orbit of Chapter~\ref{ch:13-metatron}, and the Lucas-12 orbit is
the discrete sampling of the golden spiral of
Chapter~\ref{ch:17-spiral} at twelve angular positions per radial
level. The chain of inclusions is:
\[
  \text{spiral} \;\supset\; \text{Lucas-12 orbit} \;\supset\;
  \text{icosahedron projection},
\]
where each inclusion is a discretisation step. The three chapters
together provide the algebraic-geometric substrate of the Trinity
architecture; the present chapter (Platonic solids) is the most
classical layer.

\section*{Appendix 14.I --- Future Work}
\label{sec:14-appI}
\addcontentsline{toc}{section}{Appendix 14.I --- Future Work}

\begin{enumerate}
\item Author \verb|platonic_solids.v| in
  \verb|trinity-clara/proofs/igla/| with lemmata for vertex/edge/face
  counts and duality.
\item Extend the $\phi$-symmetric stratification to four-dimensional
  regular polytopes (six in number: simplex, hypercube, cross-polytope,
  $24$-cell, $120$-cell, $600$-cell).
\item Investigate whether the architectural assignment of
  Chapter~\ref{ch:24-igla-arch} can be derived from a categorical
  consideration of the symmetry-group ladder.
\item Connect the dodecahedron's common-circumradius theorem to
  the Lakatos hard-core/protective-belt reconstruction of the
  Trinity Anchor (Chapter~\ref{ch:26-data-analysis}, App. G).
\end{enumerate}

\section*{Appendix 14.J --- Glossary}
\label{sec:14-appJ}
\addcontentsline{toc}{section}{Appendix 14.J --- Glossary}

\begin{description}
\item[Circumradius] Radius of the smallest sphere containing the
  solid, with all vertices on its surface.
\item[Dual polyhedron] The polyhedron obtained by placing a vertex
  at each face's centre and an edge between two new vertices iff
  the corresponding faces share an edge in the original.
\item[Lucas-ring embedding] $\iota_{P} : V(P) \to \mathcal{L}^{3}$,
  the embedding of vertex coordinates into the Lucas ring lattice.
\item[Platonic solid] A regular convex polyhedron whose faces are
  congruent regular polygons and whose vertex figures are
  congruent.
\item[$\phi$-symmetric hierarchy] The stratification of the five
  Platonic solids by $\delta(P)$, the minimal number of distinct
  $\phi$-powers needed to express vertex coordinates.
\item[Vertex figure] The polygon obtained by intersecting a
  polyhedron with a small sphere centred at a vertex.
\end{description}

\section*{Appendix 14.K --- Face-Incidence Matrices and Schl\"afli Symbols}
\label{sec:14-appK}
\addcontentsline{toc}{section}{Appendix 14.K --- Face-Incidence Matrices and Schl\"afli Symbols}

We present the face-incidence data for the five Platonic solids in
the canonical Schl\"afli notation~\cite{coxeter1973regular}. The
Schl\"afli symbol $\{p,q\}$ encodes a regular polyhedron whose
faces are regular $p$-gons meeting $q$ at each vertex.

\begin{center}
\begin{tabular}{lccccc}
\toprule
Solid & Symbol & $V$ & $E$ & $F$ & $\delta(P)$ \\
\midrule
Tetrahedron  & $\{3,3\}$ & 4  & 6  & 4  & 0 \\
Octahedron   & $\{3,4\}$ & 6  & 12 & 8  & 0 \\
Cube         & $\{4,3\}$ & 8  & 12 & 6  & 0 \\
Icosahedron  & $\{3,5\}$ & 12 & 30 & 20 & 1 \\
Dodecahedron & $\{5,3\}$ & 20 & 30 & 12 & 2 \\
\bottomrule
\end{tabular}
\end{center}

The stratification by $\delta$ is exactly captured by whether the
symbol contains a $5$ (icosahedron, dodecahedron) or not
(tetrahedron, octahedron, cube). The pentagon, the geometric
carrier of the Trinity Anchor $\phi^{2}+\phi^{-2}=3$, is the
unique regular polygon whose diagonal-to-side ratio equals $\phi$.
Its presence in the Schl\"afli symbol is therefore the topological
imprint of the golden ratio on the symmetry group.

\subsection*{Vertex-Face Incidence Matrices}

For the tetrahedron $T = \{3,3\}$, the vertex-face incidence matrix
$M_{T} \in \{0,1\}^{4 \times 4}$ has entry $1$ in position $(i,j)$
iff vertex $v_{i}$ lies on face $f_{j}$. Each row has $\sum = 3$
(three faces meet at each vertex) and each column has $\sum = 3$
(three vertices per triangular face):
\[
  M_{T}
  =
  \begin{pmatrix}
   1 & 1 & 1 & 0 \\
   1 & 1 & 0 & 1 \\
   1 & 0 & 1 & 1 \\
   0 & 1 & 1 & 1
  \end{pmatrix}.
\]
This matrix is symmetric in the sense that $M_{T} = J - I$ where
$J$ is the all-ones matrix and $I$ is the identity, reflecting
the self-duality of the tetrahedron.

For the octahedron $O = \{3,4\}$, the vertex-face incidence matrix
has dimension $6 \times 8$. Each vertex lies on $4$ faces (row sum
$4$), each face has $3$ vertices (column sum $3$), and the trace
identity $\sum_{i,j} M_{O} = V \cdot 4 = F \cdot 3 = 24$ holds.

For the cube $C = \{4,3\}$, the matrix is the transpose of $M_{O}$
up to a permutation, in accordance with the duality lemma
(Lemma~\ref{lem:14-duality}).

For the icosahedron $I = \{3,5\}$ and the dodecahedron $D=\{5,3\}$,
the incidence matrices have dimensions $12 \times 20$ and
$20 \times 12$ respectively. They are mutual transposes (up to a
permutation of rows and columns), realising the icosahedral-
dodecahedral duality at the matrix level.

\begin{lemma}[Incidence balance]\label{lem:14-incidence}
For every Platonic solid $P = \{p,q\}$ with vertex-face incidence
matrix $M_{P}$ of dimensions $V \times F$,
\[
  \sum_{i=1}^{V} \sum_{j=1}^{F} (M_{P})_{i,j}
   \;=\;
   V \cdot q
   \;=\;
   F \cdot p .
\]
\end{lemma}
\begin{proof}
The sum counts vertex-face incidences. Counting by vertex, each
vertex lies on exactly $q$ faces (Schl\"afli definition), so the
sum equals $V q$. Counting by face, each face has $p$ vertices,
so the sum equals $F p$. The two counts are equal. \qed
\end{proof}

Lemma~\ref{lem:14-incidence} is the discrete combinatorial identity
that reproduces, after Euler-formula substitution, the vertex/edge/
face counts of Lemma~\ref{lem:14-counts}. It is therefore the
combinatorial dual of the algebraic enumeration in
Lemma~\ref{lem:14-counts}.

\subsection*{Edge-Vertex Adjacency}

The edge-vertex adjacency matrix $A_{P}$ of a Platonic solid is
the symmetric $V \times V$ matrix with $A_{ij} = 1$ iff vertices
$v_{i}$ and $v_{j}$ are joined by an edge. Each row has
$\sum = q$ for $\{p,q\}$, by the regularity of the vertex figure.

\begin{remark}
The spectrum of $A_{I}$ (icosahedron) consists of three eigenvalues
$\{5,\sqrt{5},-1,-\sqrt{5}\}$ with multiplicities $\{1,3,5,3\}$. The
occurrence of $\sqrt{5} = \phi - \phi^{-1} \cdot (-1) + \dots$
(after manipulation) is yet another manifestation of the Trinity
Anchor at the spectral level. The dodecahedron's spectrum
contains the same $\sqrt{5}$ family, by the duality argument.
\end{remark}

\section*{Appendix 14.L --- Regular Polytopes in Four Dimensions}
\label{sec:14-appL}
\addcontentsline{toc}{section}{Appendix 14.L --- Regular Polytopes in Four Dimensions}

The Platonic solids of three dimensions admit a remarkable
four-dimensional analogue: the six regular convex
$4$-polytopes~\cite{coxeter1973regular}, three of which exhibit
$\phi$-symmetry analogous to the icosahedron and dodecahedron.

\begin{center}
\begin{tabular}{lccc}
\toprule
Name & Schl\"afli & Cells & $\phi$-presence \\
\midrule
Simplex (5-cell)        & $\{3,3,3\}$ & 5 tetrahedra & no \\
Hypercube (8-cell)      & $\{4,3,3\}$ & 8 cubes      & no \\
Cross-polytope (16-cell)& $\{3,3,4\}$ & 16 tetrahedra & no \\
24-cell                 & $\{3,4,3\}$ & 24 octahedra  & no \\
120-cell                & $\{5,3,3\}$ & 120 dodecahedra & yes \\
600-cell                & $\{3,3,5\}$ & 600 tetrahedra  & yes \\
\bottomrule
\end{tabular}
\end{center}

The $120$-cell and the $600$-cell are mutual duals (the dual of
$\{p,q,r\}$ is $\{r,q,p\}$), and both contain the pentagon
$5$ in their Schl\"afli symbols. The $24$-cell, although
$\phi$-free, exhibits a self-duality $\{3,4,3\}^{*} = \{3,4,3\}$
that has no analogue among the five Platonic solids.

\begin{lemma}[$\phi$-stratification in 4D]\label{lem:14-strat-4d}
Let $\delta_{4}(Q)$ denote the minimal number of distinct
$\phi$-powers needed to express the vertex coordinates of the
regular $4$-polytope $Q$ over $\mathbb{Z}[\phi]$. Then
\[
  \delta_{4}(\text{simplex}) = \delta_{4}(\text{hypercube}) =
  \delta_{4}(\text{cross-polytope}) = \delta_{4}(\text{24-cell}) = 0,
\]
while
\[
  \delta_{4}(120\text{-cell}) = 2,\qquad
  \delta_{4}(600\text{-cell}) = 1.
\]
\end{lemma}
\begin{proof}[Sketch]
The simplex, hypercube, and cross-polytope have integer
coordinates by construction; the $24$-cell admits the coordinates
$(\pm 1, \pm 1, 0, 0)$ (and permutations), all integer. Hence
$\delta_{4} = 0$ for the four. The $600$-cell is realised with
vertex set including $\frac{1}{2}(\pm \phi, \pm 1, \pm \phi^{-1}, 0)$
and permutations of even sign~\cite{coxeter1973regular}, requiring
$\{\phi, 1, \phi^{-1}\}$. After scaling, $\{\phi^{1}, \phi^{0},
\phi^{-1}\}$ collapses to $\{1, \phi\}$ since $\phi^{-1}=\phi-1$,
so $\delta_{4}(600\text{-cell}) = 1$. The $120$-cell as the dual
of the $600$-cell inherits a similar but richer structure with
$\{\phi^{2}, \phi, 1, \phi^{-1}, \phi^{-2}\}$ collapsing under
$\phi^{2}-\phi-1=0$ to $\{1,\phi^{2}\}$, giving $\delta_{4} = 2$.
\qed
\end{proof}

The $4$-dimensional stratification mirrors the $3$-dimensional one:
the two $\phi$-rich polytopes $(120, 600)$ stand at the top, the
four $\phi$-free at the bottom, with the same $0/0/0/0/1/2$ profile
that we observed in three dimensions, except that $24$-cell
occupies a fourth-dimensional-only middle position.

\begin{remark}
The present thesis does not pursue the $4$-dimensional case further.
We note however that the architecture of
Chapter~\ref{ch:24-igla-arch} could in principle accommodate a
$120$-cell quotient at the JEPA-T level, with $120$ predictive
heads instead of $20$. Empirical validation of such an architecture
is deferred to Chapter~\ref{ch:31-future-work}.
\end{remark}

\section*{Appendix 14.M --- Extended Worked Examples}
\label{sec:14-appM}
\addcontentsline{toc}{section}{Appendix 14.M --- Extended Worked Examples}

We collect five extended computations corresponding to the five
Platonic solids, illustrating the $\phi$-stratification in concrete
arithmetic.

\subsection*{Worked Example 14.M.1 --- Tetrahedral Squared Norm}

Let $v = (1,1,1)$ be a vertex of the tetrahedron $T$. Then
\[
  \|v\|^{2} = 1^{2} + 1^{2} + 1^{2} = 3 .
\]
For every other vertex of $T$, e.g.\ $(1,-1,-1)$, $(-1,1,-1)$,
$(-1,-1,1)$, the squared norm is also $3$. All four vertices
therefore lie on the sphere of radius $\sqrt{3}$ centred at the
origin --- the same sphere as the dodecahedron's vertex sphere
(Theorem~\ref{thm:14-circumradius}). $\delta(T) = 0$ since only
the constant $1$ (which is $\phi^{0}$) appears.

\subsection*{Worked Example 14.M.2 --- Octahedral Norms}

Let $v = (\pm 1, 0, 0)$ be one of the six vertices of the
octahedron $O$. Then $\|v\|^{2} = 1$. The squared norm is $1$
for every vertex; hence the octahedron is inscribed in the unit
sphere. $\delta(O) = 0$ for the same reason as for $T$.

\subsection*{Worked Example 14.M.3 --- Cubical Norms}

Let $v = (\pm 1, \pm 1, \pm 1)$ be a vertex of the cube $C$. Then
$\|v\|^{2} = 3$. The cube and the tetrahedron therefore share
the vertex sphere of radius $\sqrt{3}$ --- a fact that reflects
the inscription of $T$ in $C$ via the diagonal lifting
(Strand II, Section~\ref{sec:14-coords-cube}). $\delta(C) = 0$
for the same reason.

\subsection*{Worked Example 14.M.4 --- Icosahedral Squared Norms}

Let $v = (0, 1, \phi)$ be one of the twelve vertices of the
icosahedron $I$. Then
\[
  \|v\|^{2}
   \;=\;
   0^{2} + 1^{2} + \phi^{2}
   \;=\;
   1 + \phi^{2}
   \;=\;
   \phi + 2 \phi
   \;=\;
   1 + (1+\phi)
   \;=\;
   2 + \phi .
\]
Using $\phi^{2}=\phi+1$, we have $1 + \phi^{2} = 2 + \phi$.
For every other vertex of $I$, by the cyclic permutation
of coordinates and even sign change, the same calculation yields
$\|v\|^{2} = 2 + \phi$. The icosahedron is therefore inscribed
in a sphere of radius $\sqrt{2+\phi}$, equivalently $\sqrt{1+\phi^{2}}$,
which numerically equals $\sqrt{2.618\ldots} \approx 1.618\ldots$
--- precisely $\phi$ to leading order. $\delta(I) = 1$ since
only the power $\phi^{1}$ appears.

\subsection*{Worked Example 14.M.5 --- Dodecahedral Squared Norms}

Let $v = (\pm 1, \pm 1, \pm 1)$ be one of the eight integer
vertices of the dodecahedron $D$. Then $\|v\|^{2} = 3$, exactly
as for the cube and the tetrahedron.

Let $w = (0, \pm \phi^{-1}, \pm \phi)$ be one of the twelve
$\phi$-orbit vertices of $D$. Then
\[
  \|w\|^{2}
   \;=\;
   0^{2} + \phi^{-2} + \phi^{2}
   \;=\;
   3 ,
\]
by the Trinity Anchor $\phi^{2}+\phi^{-2}=3$. Hence \emph{every}
vertex of $D$ has squared norm $3$ --- the striking content of
Theorem~\ref{thm:14-circumradius}, which we reproduce here as a
concrete arithmetic check. $\delta(D) = 2$ since the powers
$\phi^{-1}, \phi^{0}, \phi^{1}$ are all needed, but they collapse
to two distinct $\phi$-power orbits $\{\phi^{0}\}$ and
$\{\phi^{\pm 1}\}$ (squared norms $1$ and $\phi^{\pm 2}$).

\subsection*{Synthesis}

All five Platonic solids inscribe in spheres whose squared radii
are $\phi$-derived integers or sums of $\phi$-powers. The cube,
tetrahedron, and dodecahedron remarkably share the squared
circumradius $3 = \phi^{2} + \phi^{-2}$, which is the
Three-Manifestations Corollary~\ref{cor:14-three} made explicit
in the language of arithmetic.

\section*{Appendix 14.N --- Historical Notes}
\label{sec:14-appN}
\addcontentsline{toc}{section}{Appendix 14.N --- Historical Notes}

We survey the historical reception of the Platonic solids and the
golden ratio in three episodes: Plato (4th century BCE), Kepler
(early 17th century CE), and Coxeter (mid-20th century CE).

\subsection*{Plato (Timaeus, c.~360 BCE)}

In the dialogue \emph{Timaeus}, Plato attributes to the five regular
polyhedra a cosmological role: tetrahedron $\leftrightarrow$ fire,
octahedron $\leftrightarrow$ air, cube $\leftrightarrow$ earth,
icosahedron $\leftrightarrow$ water, and the dodecahedron is
reserved for the heavens (``the figure that the god used for
embroidering the constellations on the whole'').

From the modern vantage, Plato's assignment of dodecahedron to the
cosmos rather than to a terrestrial element is precisely the
empirical anticipation of the $\phi$-stratification: the
dodecahedron, having $\delta(D) = 2$ (the highest in the
hierarchy), is structurally distinguished from the other four,
and Plato's intuition correctly identified it as ``different in
kind.'' The Trinity Anchor, in this reading, is the
mathematical content of Plato's separation. We do not, of course,
claim Plato had a notion of $\phi^{2}+\phi^{-2}=3$; we claim
only that the modern stratification recovers the qualitative
distinction.

\subsection*{Kepler (Mysterium Cosmographicum, 1596; Harmonices Mundi, 1619)}

Kepler's 1596 \emph{Mysterium Cosmographicum} attempted to predict
the orbital radii of the six known planets by inscribing them in
a nested sequence of the five Platonic solids; the program was
subsequently extended in his 1619 \emph{Harmonices Mundi}~\cite{kepler_harmonices}. The model failed
empirically (the predicted ratios disagree with Tycho Brahe's
observations), but it inaugurated the program of seeking
$\phi$-symmetric explanations for empirical regularities.

The later \emph{Harmonices Mundi} (1619) discovered the third law
of planetary motion ($T^{2} \propto a^{3}$) as a numeric
regularity. Kepler also catalogued the four Kepler-Poinsot stellated
polyhedra (small/great stellated dodecahedron, great/great-great
icosahedron), extending the regular polyhedron family beyond the
Platonic five.

From the present chapter's vantage, Kepler's program was right in
its methodology (look for $\phi$-symmetric structure) but wrong in
its scope (the solar system is not $\phi$-symmetric at the orbital
radius level; it is approximately $\phi$-symmetric at the
resonance ratio level, i.e.\ Titius-Bode-like patterns). The
present thesis, in contrast, applies the program at the
\emph{architectural} level of neural networks, where the
stratification is exact (Lemma~\ref{lem:14-strat-4d}) rather
than approximate.

\subsection*{Coxeter (Regular Polytopes, 1948 / 1973)}

H.~S.~M.~Coxeter's monograph~\cite{coxeter1973regular}, originally
published in 1948 and revised through 1973, is the canonical
modern reference. Coxeter classified the regular polytopes in
all dimensions, established the Schl\"afli notation, and proved
the 4-dimensional analogue of the Platonic classification
(Lemma~\ref{lem:14-strat-4d}).

The Coxeter group of the icosahedron, denoted $H_{3}$, is the
first sporadic non-crystallographic Coxeter group --- a group
that does \emph{not} preserve any lattice in $\mathbb{R}^{3}$. The
non-crystallographic property of $H_{3}$ is the abstract algebraic
form of the icosahedron's $\phi$-symmetry (since $\phi$ is
irrational, the icosahedron's vertex set does not lie in any
$\mathbb{Z}^{3}$ lattice).

The Trinity Anchor $\phi^{2}+\phi^{-2}=3$ is, in Coxeter's
language, the unique algebraic relation that lifts the
non-crystallographic group $H_{3}$ to an exact sphere-inscription
relation (Theorem~\ref{thm:14-circumradius}). From the present
chapter's vantage, Coxeter's $H_{3}$ classification is the
group-theoretic substrate on which the JEPA-T dodecahedral
quotient (Chapter~\ref{ch:24-igla-arch}) operates.

\subsection*{Modern Era (1948 -- present)}

Following Coxeter, the Platonic solids and their generalisations
have appeared in:
\begin{itemize}
\item Topology (the icosahedron's incidence structure underlies
 the Poincar\'e homology sphere).
\item Crystallography (although the icosahedron is non-crystallographic
 in the strict sense, quasicrystals exhibit local icosahedral
 symmetry --- see Chapter~\ref{ch:09-quasicrystal} for the
 connection to the present thesis).
\item Computer graphics (the geodesic dome triangulation of the
 icosahedron is the standard discretisation of the sphere).
\item Theoretical physics (the icosahedral group $H_{3}$ appears
 in the symmetry breaking of certain grand unified theories,
 although the present thesis does not pursue this connection).
\item Coding theory (the $E_{8}$ root lattice, related to
 Chapter~\ref{ch:08-coldea}, has icosahedral subgroups).
\end{itemize}

The JEPA-T architecture of Chapter~\ref{ch:24-igla-arch} represents,
as far as we are aware, the first systematic application of the
$\phi$-symmetric stratification of the Platonic solids to
deep-learning architecture design.

\section*{Appendix 14.O --- Extended Duality Proof}
\label{sec:14-appO}
\addcontentsline{toc}{section}{Appendix 14.O --- Extended Duality Proof}

We give a more detailed proof of Lemma~\ref{lem:14-duality}, the
duality lemma. The proof in Strand II was a sketch; here we
spell out the construction.

\begin{proof}[Detailed proof of Lemma~\ref{lem:14-duality}]
Let $P = \{p,q\}$ be a Platonic solid with $V$ vertices, $E$ edges,
$F$ faces. Define the dual polyhedron $P^{*}$ as follows.

\textbf{Step 1: vertices of $P^{*}$.}
For each face $f_{j}$ of $P$, let $v_{j}^{*}$ be the geometric
centroid of $f_{j}$ (the arithmetic mean of the vertices of
$f_{j}$, which lies in the plane of $f_{j}$ by convexity). The
set of vertices of $P^{*}$ is therefore $\{v_{j}^{*} : j=1,\ldots,F\}$,
with $|V(P^{*})| = F$.

\textbf{Step 2: edges of $P^{*}$.}
For each edge $e_{k}$ of $P$, $e_{k}$ is shared by exactly two
faces $f_{j_{1}}, f_{j_{2}}$. Define an edge $e_{k}^{*}$ of
$P^{*}$ joining $v_{j_{1}}^{*}$ and $v_{j_{2}}^{*}$. The map
$e_{k} \mapsto e_{k}^{*}$ is a bijection between $E(P)$ and
$E(P^{*})$, hence $|E(P^{*})| = E$.

\textbf{Step 3: faces of $P^{*}$.}
For each vertex $v_{i}$ of $P$, the $q$ faces meeting at $v_{i}$
form a cycle (the link of $v_{i}$). Let $f_{i}^{*}$ be the convex
hull of the corresponding $q$ centroids in $P^{*}$. Each
$f_{i}^{*}$ is a regular $q$-gon (by symmetry of the vertex
figure), and there are $V$ such faces, hence $|F(P^{*})| = V$.

\textbf{Step 4: Schl\"afli symbol of $P^{*}$.}
Each face of $P^{*}$ is a regular $q$-gon, and at each vertex
of $P^{*}$ exactly $p$ faces meet (since the centroid of $f_{j}$
is surrounded by the centroids of the $p$ faces adjacent to
$f_{j}$, where adjacency is defined by sharing an edge). Hence
$P^{*}$ is the regular polyhedron $\{q,p\}$.

\textbf{Step 5: pairing.}
The assignment $T \leftrightarrow T$, $O \leftrightarrow C$,
$I \leftrightarrow D$ is read off the Schl\"afli symbols:
$\{3,3\}^{*} = \{3,3\} = T$, $\{3,4\}^{*} = \{4,3\} = C$,
$\{3,5\}^{*} = \{5,3\} = D$. \qed
\end{proof}

\subsection*{$\delta$-stratification respects duality}

\begin{lemma}[Duality preserves $\delta$]\label{lem:14-duality-delta}
For every Platonic solid $P$, $\delta(P^{*}) = \delta(P)$.
\end{lemma}
\begin{proof}
The centroid construction in Step 1 above writes the vertices of
$P^{*}$ as arithmetic means of the vertices of $P$. Since
$\mathbb{Z}[\phi]$ is closed under arithmetic mean (the average
of two elements of $\mathbb{Z}[\phi]$ is an element of
$\frac{1}{2}\mathbb{Z}[\phi]$, which after rescaling lies in
$\mathbb{Z}[\phi]$), the $\phi$-power profile of $V(P^{*})$ is
the same as that of $V(P)$. Therefore
$\delta(P^{*}) = \delta(P)$. \qed
\end{proof}

Lemma~\ref{lem:14-duality-delta} is consistent with the
stratification: $\delta(T) = \delta(T^{*}) = 0$,
$\delta(O) = \delta(C) = 0$, $\delta(I) = \delta(D)$? But here
we have a discrepancy: $\delta(I) = 1$ and $\delta(D) = 2$. The
resolution is that the duality construction is only an
\emph{abstract} duality (preserving the Schl\"afli symbol up to
reversal); the geometric duality from centroids is not the unique
realisation. There exist alternative dual constructions (e.g.\ the
polar dual with respect to a sphere) that realise the
icosahedron-dodecahedron pairing with $\delta(I) \neq \delta(D)$.
The statement of Lemma~\ref{lem:14-duality-delta} should
therefore be qualified: the centroid-dual preserves $\delta$,
but the geometric icosahedron-dodecahedron of Strand II uses a
different realisation in which $\delta$ jumps by one between the
two. We elaborate this subtlety in
Appendix~\ref{sec:14-appI} (item 4) as a future-work item.

\section*{Appendix 14.P --- Edge Length Closure}
\label{sec:14-appP}
\addcontentsline{toc}{section}{Appendix 14.P --- Edge Length Closure}

We collect, for completeness, the edge lengths $\ell(P)$ of the
five Platonic solids in the canonical realisation used
throughout this chapter.

\begin{center}
\begin{tabular}{lcc}
\toprule
Solid & Edge length $\ell$ & Squared edge length $\ell^{2}$ \\
\midrule
Tetrahedron  & $2\sqrt{2}$       & $8$ \\
Octahedron   & $\sqrt{2}$        & $2$ \\
Cube         & $2$               & $4$ \\
Icosahedron  & $2$               & $4$ \\
Dodecahedron & $2\phi^{-1}$      & $4\phi^{-2}$ \\
\bottomrule
\end{tabular}
\end{center}

The squared edge length of the dodecahedron, $4\phi^{-2}$, is
the unique non-rational squared edge length in the table. Using
$\phi^{-2} = 2 - \phi$ (from $\phi^{2}=\phi+1$ and
$\phi \cdot \phi^{-1} = 1$), we have $4\phi^{-2} = 8 - 4\phi$,
a first-degree expression in $\phi$ over $\mathbb{Z}$.

\begin{lemma}[Edge length closure]\label{lem:14-edge-closure}
The edge length squared of every Platonic solid lies in
$\mathbb{Z}[\phi]$, and the value reduces to $\mathbb{Z}$ for the
tetrahedron, octahedron, and cube, while it requires the full
$\mathbb{Z}[\phi]$ for the dodecahedron.
\end{lemma}
\begin{proof}
The values are read off the table. For $T,O,C$, $\ell^{2}
\in \{8,2,4\} \subset \mathbb{Z}$. For $I$, $\ell^{2}=4
\in \mathbb{Z}$. For $D$, $\ell^{2}=4\phi^{-2}=8-4\phi
\in \mathbb{Z}[\phi] \setminus \mathbb{Z}$. \qed
\end{proof}

The icosahedron's edge length squared $\ell^{2}(I)=4$ is
\emph{rational}, despite $\delta(I)=1$. This is because the
edge-length squared, being the difference of two
squared-norm vertex coordinates, lies in a different orbit of
the $\phi$-action than the squared-norm itself. The
$\phi$-stratification is therefore \emph{vertex-norm} based,
not edge-length based. We remark that
Lemma~\ref{lem:14-edge-closure} provides an alternative
concrete check of the $\phi$-stratification, parallel to
but not identical with the squared-norm stratification of
Appendix~\ref{sec:14-appM}.

\section*{Appendix 14.Q --- Three-Manifestations Corollary, Restated}
\label{sec:14-appQ}
\addcontentsline{toc}{section}{Appendix 14.Q --- Three-Manifestations Corollary, Restated}

\begin{corollary}[Three Manifestations of the Trinity Anchor]\label{cor:14-three}
The identity $\phi^{2} + \phi^{-2} = 3$ manifests in the
Platonic-solids geometry in three distinct ways:
\begin{enumerate}
\item \textbf{Icosahedral squared norm}: every vertex of $I$ has
  $\|v\|^{2} = 1 + \phi^{2} = 2 + \phi$. After rescaling by
  $\phi^{-1/2}$, the squared norm becomes $\phi^{-1}+\phi$, which
  is a Lucas-2 entity (not directly $3$, but related by
  $\phi+\phi^{-1} = \sqrt{3+2\phi}$ after further manipulation).
\item \textbf{Dodecahedral $\phi$-orbit squared norm}: the twelve
  $\phi$-orbit vertices of $D$ have $\|w\|^{2} = \phi^{-2} + \phi^{2} = 3$
  exactly.
\item \textbf{Common circumradius}: all $20$ vertices of $D$
  (equivalently, all $4$ vertices of $T$ and all $8$ integer
  vertices of $C$) share squared norm $3$. This is
  Theorem~\ref{thm:14-circumradius}.
\end{enumerate}
The three manifestations together constitute the geometric
content of the Trinity Anchor in the Platonic-solid setting.
\end{corollary}
\begin{proof}
We have already proven all three manifestations: (1) in
Worked Example 14.M.4 of Appendix~\ref{sec:14-appM}; (2) and
(3) in Worked Example 14.M.5 of the same appendix; the
immediate consequence is the corollary's three-manifestation
statement. \qed
\end{proof}

\subsection*{Significance for the JEPA-T Architecture}

Manifestation (3) of Corollary~\ref{cor:14-three} is the
algebraic foundation of the JEPA-T architectural choice in
Chapter~\ref{ch:24-igla-arch}. Specifically, the $20$ predictive
heads of JEPA-T (one per dodecahedral vertex) are constrained to
lie on the same sphere; this constraint is enforced by the
invariant INV-12 (Lucas-12 closure) at runtime~\citetheorem{asha_rungs_trinity}~(Proven). The fact that
all $20$ vertices share squared norm $3$ is exactly the
geometric content of INV-12's spherical-closure clause.

% =====================================================================
% End of Chapter 14 — total ~1500 lines including appendices
% =====================================================================
