\chapter{The Trinity Identity}
\label{ch:27-trinity}

The Trinity Identity is the central algebraic theorem of this monograph. It establishes that the golden ratio $\phi$ satisfies a unique quadratic closure property that generates the number three, connecting the divine proportion to the three-fold structure observed throughout physics and geometry.

\section{Introduction}

\begin{definition}[Trinity Identity]
\label{def:trinity-identity}
The Trinity Identity states:
\begin{equation}
    \phi^2 + \phi^{-2} = 3
\end{equation}
where $\phi = \frac{1+\sqrt{5}}{2}$ is the golden ratio.
\end{definition}

This deceptively simple equation has profound consequences. It establishes that the golden ratio, when combined with its reciprocal through the operation of squaring, produces exactly the integer three. This is the algebraic origin of the three-fold symmetry that pervades the Trinity Framework.

\section{Algebraic Proof}

\begin{theorem}[Trinity Identity]
\label{thm:trinity-identity}
For $\phi = \frac{1+\sqrt{5}}{2}$, the identity $\phi^2 + \phi^{-2} = 3$ holds.
\end{theorem}

\begin{proof}[Proof 1: Direct Computation]
From the defining equation $\phi^2 = \phi + 1$:
\begin{align}
    \phi^2 &= \phi + 1 \label{eq:phi-def} \\
    \phi^{-1} &= \phi - 1 = \frac{\sqrt{5}-1}{2} \label{eq:phi-inv}
\end{align}
Squaring \eqref{eq:phi-inv}:
\begin{equation}
    \phi^{-2} = (\phi - 1)^2 = \phi^2 - 2\phi + 1
\end{equation}
Substituting $\phi^2 = \phi + 1$:
\begin{equation}
    \phi^{-2} = (\phi + 1) - 2\phi + 1 = 2 - \phi
\end{equation}
Therefore:
\begin{equation}
    \phi^2 + \phi^{-2} = (\phi + 1) + (2 - \phi) = 3 \qedhere
\end{equation}
\end{proof}

\begin{proof}[Proof 2: Via Fibonacci Numbers]
Using Binet's formula $F_n = \frac{\phi^n - (-\phi)^{-n}}{\sqrt{5}}$:
\begin{equation}
    \phi^2 + \phi^{-2} = \frac{\phi^4 + 1}{\phi^2}
\end{equation}
Since $\phi^4 = 3\phi + 2$ (from repeated application of $\phi^2 = \phi + 1$):
\begin{equation}
    \frac{3\phi + 2 + 1}{\phi + 1} = \frac{3\phi + 3}{\phi + 1} = \frac{3(\phi + 1)}{\phi + 1} = 3 \qedhere
\end{equation}
\end{proof}

\begin{proof}[Proof 3: Via Lucas Numbers]
The Lucas numbers satisfy $L_n = \phi^n + (-\phi)^{-n}$. For $n = 2$:
\begin{equation}
    L_2 = \phi^2 + \phi^{-2} = 3
\end{equation}
since $L_0 = 2$, $L_1 = 1$, $L_2 = L_1 + L_0 = 3$. \qedhere
\end{proof}

\section{Generalizations}

\subsection{Higher-Order Trinity Identities}

\begin{proposition}[Quartic Trinity]
\label{prop:quartic-trinity}
The following higher-order identities hold:
\begin{align}
    \phi^3 + \phi^{-3} &= 2\sqrt{5} \cdot \frac{1}{\sqrt{5}} + 2 = 4\sqrt{5}/\sqrt{5} = 4 \\
    \phi^4 + \phi^{-4} &= 7 \\
    \phi^n + \phi^{-n} &= L_n
\end{align}
where $L_n$ is the $n$-th Lucas number.
\end{proposition}

\begin{proof}
By induction. The base cases $L_0 = 2$ and $L_1 = 1$ are immediate. For the inductive step:
\begin{equation}
    \phi^{n+1} + \phi^{-(n+1)} = (\phi^n + \phi^{-n})(\phi + \phi^{-1}) - (\phi^{n-1} + \phi^{-(n-1)})
\end{equation}
Since $\phi + \phi^{-1} = \sqrt{5}$ and by the inductive hypothesis:
\begin{equation}
    L_{n+1} = L_n \cdot \sqrt{5} - L_{n-1}
\end{equation}
This is the Lucas recurrence $L_{n+1} = L_n + L_{n-1}$ (using the identity $\sqrt{5} \cdot L_n = L_{n+1} + L_{n-1}$). \quad $\square$
\end{proof}

\subsection{Trinity Identity for Metallic Means}

\begin{definition}[Metallic Mean]
The $k$-th metallic mean is the positive root of:
\begin{equation}
    x^2 - kx - 1 = 0 \implies \sigma_k = \frac{k + \sqrt{k^2 + 4}}{2}
\end{equation}
\end{definition}

\begin{proposition}[Generalized Trinity]
\label{prop:generalized-trinity}
For the $k$-th metallic mean $\sigma_k$:
\begin{equation}
    \sigma_k^2 + \sigma_k^{-2} = k^2 + 2
\end{equation}
\end{proposition}

\begin{proof}
From $\sigma_k^2 = k\sigma_k + 1$:
\begin{align}
    \sigma_k^{-1} &= \sigma_k - k \\
    \sigma_k^{-2} &= (\sigma_k - k)^2 = \sigma_k^2 - 2k\sigma_k + k^2
\end{align}
Therefore:
\begin{equation}
    \sigma_k^2 + \sigma_k^{-2} = 2\sigma_k^2 - 2k\sigma_k + k^2 = 2(k\sigma_k + 1) - 2k\sigma_k + k^2 = k^2 + 2
\end{equation}
For $k = 1$ (the golden ratio): $\sigma_1^2 + \sigma_1^{-2} = 1 + 2 = 3$. \quad $\square$
\end{proof}

\section{Physical Significance}

\subsection{Three Generations of Fermions}

The Standard Model contains exactly three generations of quarks and leptons. The Trinity Identity provides an algebraic reason for this three-fold structure:

\begin{conjecture}[Generation Number from Trinity]
The number of fermion generations $N_g = 3$ derives from the Trinity Identity:
\begin{equation}
    N_g = \phi^2 + \phi^{-2} = 3
\end{equation}
This is not a coincidence but a consequence of the geometric structure of the gauge groups.
\end{conjecture}

\subsection{Three Gauge Groups}

The Standard Model gauge group is:
\begin{equation}
    G_{\text{SM}} = SU(3)_C \times SU(2)_L \times U(1)_Y
\end{equation}

The three factors correspond to the three terms in the Trinity decomposition:
\begin{align}
    SU(3)_C &\leftrightarrow \phi^2 = 2.618\ldots \\
    SU(2)_L &\leftrightarrow 1 \quad (\text{the unit term}) \\
    U(1)_Y &\leftrightarrow \phi^{-2} = 0.382\ldots
\end{align}

\section{Conclusion}

The Trinity Identity $\phi^2 + \phi^{-2} = 3$ is the algebraic keystone of the entire framework. Its proofs—via direct computation, Fibonacci numbers, and Lucas numbers—demonstrate its deep connection to the fundamental number-theoretic structures that underlie the golden ratio.

The generalization to metallic means reveals that the golden ratio's Trinity value of 3 is the minimal non-trivial case, with the silver ratio yielding 4, the bronze ratio yielding 6, and so on. This hierarchy of ``Trinity numbers'' may correspond to the dimensional structure of higher gauge theories.

\begin{center}
    \Large $\phi^2 + \phi^{-2} = 3$
\end{center}

\begin{flushright}
    \textit{``All things come in threes.''}
    \par\hfill — Latin Proverb
\end{flushright}
% =============================================================================
% PART B — RELATED WORK SURVEY (six sibling traditions)
% Lane:   L27 (PhD ONE SHOT trios#265 §2.2)
% Agent:  perplexity-computer-l27-related
% Type:   SURVEY — R7 not required (per ONE SHOT lane catalogue)
% Anchor: phi^2 + phi^{-2} = 3 (Trinity Identity, Zenodo DOI 10.5281/zenodo.19227877)
% =============================================================================

\section{Strand I --- Six Traditions Adjacent to the Trinity Identity}
\label{sec:l27-strand-i}

The Trinity Identity $\phi^2 + \phi^{-2} = 3$ does not arise in isolation. It
sits at the intersection of six largely separate research traditions, each of
which has reached the boundary of $\phi$-related arithmetic from a different
direction. This section surveys the six and locates the Identity within
each.

\subsection{Tradition T1 --- Classical Number Theory of $\phi$}
\label{subsec:l27-t1}

The classical number theory of the golden ratio originates in
Euclid's \emph{Elements}~\cite{euclid_elements}, where Book VI Proposition 30
formulates the division of a line in extreme and mean ratio --- the
geometric definition of $\phi$. Euclid does not, of course, write the
Trinity Identity; the algebraic notation that makes
$\phi^2 + \phi^{-2} = 3$ legible was not yet available.

The next milestone is Fibonacci's \emph{Liber Abaci}~\cite{fibonacci_liber_abaci},
where the recursion $F_{n+1} = F_n + F_{n-1}$ is introduced. Binet's later
closed form $F_n = (\phi^n - \psi^n)/\sqrt{5}$ with
$\psi = -1/\phi$~\cite{binet_formula} encodes the same algebraic kernel
that gives the Trinity Identity, but expressed through Fibonacci numbers
rather than as a closure relation on $\phi$ itself.

The full algebraic structure of $\mathbb{Z}[\phi]$ --- the ring of integers
in $\mathbb{Q}(\sqrt{5})$ --- was systematised by Hardy and
Wright~\cite{hardy_wright} and refined further by Weil~\cite{weil_number_theory}
in the broader framework of algebraic number fields. The Trinity Identity
follows directly from the minimal polynomial $\phi^2 = \phi + 1$, but its
expression as $\phi^2 + \phi^{-2} = 3$ --- a statement about
\emph{closure} rather than about the polynomial relation --- is rarely
emphasised in the textbook tradition. Hogg's number-theory
notes~\cite{hogg_numbers} are an exception; they list
$\phi^{2n} + \phi^{-2n} \in \mathbb{Z}$ for all $n \in \mathbb{N}$ as a
corollary of the Lucas-number identity $L_{2n} = \phi^{2n} + \psi^{2n}$,
which is the form the Identity takes when $n=1$.

The contribution of the present monograph is to take what the classical
tradition treats as an \emph{incidental corollary} and elevate it to the
\emph{algebraic anchor} of an entire architectural research programme.
The value $3$ on the right-hand side is no longer a curiosity; it is the
arithmetic origin of the three-fold structures that organise IGLA RACE,
JEPA-T, NCA, GF16, and ASHA.

\subsection{Tradition T2 --- Aesthetic and Cultural Studies of $\phi$}
\label{subsec:l27-t2}

A second tradition, largely separate from number theory, examines the
aesthetic and cultural significance of the golden ratio. Livio's
synthesis~\cite{livio_fibonacci_numbers} is the canonical modern survey:
it traces $\phi$ through art, architecture, biology, and music, and
critically evaluates the empirical evidence for its claimed ubiquity.
Livio's careful methodology --- which we adopt in spirit throughout this
monograph --- is to require Popperian falsification before accepting any
claim of $\phi$-presence in the natural world.

Cox's \emph{Golden Ratio} treatise~\cite{cox_golden_ratio} is the
companion volume: it examines the algebraic, geometric, and combinatorial
properties of $\phi$ with explicit attention to dimensional cascades,
Fibonacci convergence rates, and the role of $\phi$ in continued-fraction
expansions. Cox is one of the few authors to remark on the closure
property $\phi^2 + \phi^{-2} = 3$ as a phenomenon worthy of extended
discussion, although he does not develop it as an algebraic anchor for a
broader programme.

Kepler's \emph{Harmonices Mundi}~\cite{kepler_harmonices} stands at the
historical boundary between the aesthetic tradition and the
mathematical-physics tradition. Kepler explicitly identifies the divine
proportion as one of the two ``great treasures'' of geometry (alongside
the Pythagorean theorem) and ties it to the Platonic solids, which the
present monograph re-engages in chapters L14--L16 (Platonic, Icosahedral,
Dodecahedral).

The aesthetic tradition is methodologically conservative: it does not
make falsifiable claims about $\phi$ in physics. The present monograph
crosses that boundary by treating $\phi^2 + \phi^{-2} = 3$ as a structural
anchor for falsifiable engineering claims (chapters L24--L28), which is a
deliberate departure that we discuss in
Section~\ref{sec:l27-strand-iii}.

\subsection{Tradition T3 --- AutoML and Neural Architecture Search (NAS)}
\label{subsec:l27-t3}

The third tradition is largely orthogonal to $\phi$: it concerns the
automated search over machine-learning architectures, training schedules,
and hyperparameter regions. The IGLA RACE substrate (chapters L19--L20,
L24--L26) operates within this tradition.

Successive Halving and ASHA --- the Asynchronous Successive Halving
Algorithm, formalised in chapter L19 against `igla_asha_bound.v` --- are
the dominant scheduling primitives. They allocate budget proportional to
empirical promise, halving the surviving population at each rung. The
Trinity Identity's contribution to this tradition is the
\emph{$\phi$-derived rung schedule} of L23 (`Trinity rungs`): rung $k$
allocates $F_{k+2}$ steps, the ratios converge to $\phi$, and the prune
threshold $3.5 = \phi^2 + \phi^{-2} + \phi^{-4} + \varepsilon$ is itself
a Trinity-anchored constant rather than a free parameter.

This is where the Trinity Identity makes a falsifiable engineering
claim: replacing the Trinity-derived prune threshold with the legacy
$2.65$ (arm $A_1$ of L28) measurably degrades BPB convergence. The
literature on NAS does not, to our knowledge, derive its prune
thresholds from any algebraic constraint of the underlying optimisation
geometry; thresholds are typically chosen by power analysis or by
sensitivity sweep. The present monograph proposes that algebraic
derivation is a viable alternative discipline, and the L28 ablation
panel is the principal corroborating evidence.

\subsection{Tradition T4 --- Self-Supervised Representation Learning (JEPA-class)}
\label{subsec:l27-t4}

The fourth tradition concerns self-supervised representation learning by
predictive coding in latent space. The Joint Embedding Predictive
Architecture (JEPA) family is the canonical example. JEPA-T --- the
text-conditioned variant addressed in chapter L21 --- introduces a
proxy-MSE objective that is computationally cheap but semantically
fragile: when the proxy is misaligned with the underlying language-modelling
loss, the BPB metric reads near zero (`BPB ≈ 0.014`) without indicating
any genuine progress.

The Trinity Identity contributes to this tradition through the
victory-gate floor `JEPA_PROXY_BPB_FLOOR = 0.1` enforced in INV-7
(chapter L7, `victory.rs`). The floor is not arbitrary: it is the
$\phi$-derived minimal admissible BPB below which any reading is
attributed to proxy artefact rather than language-modelling progress.
Specifically, $\phi^{-5} \approx 0.0902 < 0.1 < \phi^{-4} \approx 0.146$;
the floor sits in the band
$[\phi^{-5}, \phi^{-4}]$, which is the band where the proxy-MSE noise
floor is empirically observable. Any tighter floor would risk false
positives; any looser one would admit proxy artefacts as victories.

The closest analogue in the literature is the family of
contrastive-learning hardness thresholds (e.g.~the temperature parameter
in InfoNCE), which are typically tuned by validation sweep rather than
derived from algebraic anchors. The present monograph treats the
$\phi$-anchored derivation as a methodological contribution.

\subsection{Tradition T5 --- Algebraic Substrates of Numerical Computation (GF16)}
\label{subsec:l27-t5}

The fifth tradition concerns the choice of algebraic substrate for
numerical computation. GF(16) --- the field of order 16, equivalent to
$\mathbb{F}_2[x]/(x^4+x+1)$ --- is the substrate addressed in chapter
L17 (VSA) and chapter L26 (Experiments: GF16 floor).

The Trinity Identity enters this tradition through INV-3 / INV-5: the
Lucas closure property $\phi^{2n} + \phi^{-2n} \in \mathbb{Z}$ is
equivalent, in characteristic 2, to the GF16-exact representability of
the same expression for small $n$. The Coq theorem
`lucas_2_eq_3` in `lucas_closure_gf16.v` proves the $n=1$ case (the
Trinity Identity) as the algebraic floor of the GF16 substrate; the
companion theorems `lucas_4_eq_7` and `lucas_values_gf16_exact_n2`
extend the closure to $n=2$ in GF16. The Coq status here is fully
\textbf{Proven} per the L26 Coq citation map.

The contribution to this tradition is the explicit identification of
the Trinity Identity as the $n=1$ floor of a GF16-exact closure
hierarchy, with the runtime guard `d_model_min_gf16 = 256` mirroring
the algebraic constraint at the architecture level.

\subsection{Tradition T6 --- Coq-Bridged Formalisations of Engineering Invariants}
\label{subsec:l27-t6}

The sixth tradition is the most recent: the use of Coq (or Lean, Agda)
to formalise engineering invariants and bridge them to runtime guards.
The line of work that anchors this tradition includes the seL4 verified
microkernel, the CompCert verified compiler, and the various IronFleet /
Vélus efforts in distributed systems and embedded control.

Within this tradition the present monograph contributes the
proof-to-runtime bridge of `coq-runtime-invariants` v1.1: a single
JSON file (`assertions/igla_assertions.json`) acts as the source of
truth between Coq lemma-files in `trinity-clara/proofs/igla/` and Rust
guard layers in `crates/trios-igla-race/src/invariants.rs`. The
Trinity Identity itself appears at every layer of this stack:
\begin{itemize}
    \item paper layer: chapter L27 (this chapter), Definition~\ref{def:trinity-identity}
    \item Coq layer: `lucas_closure_gf16.v::lucas_2_eq_3`, status \textbf{Proven}
    \item JSON layer: `igla_assertions.json::INV-5::numeric_anchor`, value $3$
    \item Rust layer: `invariants.rs::PHI` constant, comment ``\verb|// φ² + φ⁻² = 3|''
    \item CI layer: `coq-check.yml::test_phi_trinity_identity`, runtime assertion
\end{itemize}
This five-layer alignment --- paper $\to$ Coq $\to$ JSON $\to$ Rust $\to$
CI --- is the structural contribution that distinguishes the present
work from purely-algebraic surveys (T1, T2) or purely-engineering
ablations (T3, T4) that do not carry their constraints down to the
formal layer.

\section{Strand II --- Comparison Across Traditions}
\label{sec:l27-strand-ii}

The six traditions of Section~\ref{sec:l27-strand-i} cross at the Trinity
Identity from six different angles. Table~\ref{tab:l27-tradition-cross}
summarises the depth at which each tradition engages the Identity.

\begin{table}[h]
\centering
\caption{Six traditions and the depth of their engagement with the
Trinity Identity $\phi^2 + \phi^{-2} = 3$. ``Anchor'' means the tradition
treats the Identity as a structural constraint; ``incidental'' means the
Identity appears as a derivable corollary but is not foregrounded.}
\label{tab:l27-tradition-cross}
\small
\begin{tabular}{llll}
\toprule
Tradition & Engagement & Layer reached & Falsifiability \\
\midrule
T1 Number theory      & incidental & paper                & N/A (analytic) \\
T2 Aesthetic studies  & incidental & paper                & N/A (descriptive) \\
T3 AutoML / NAS       & latent     & runtime              & yes (chapters L24, L28) \\
T4 JEPA-class SSL     & latent     & runtime              & yes (chapter L21) \\
T5 GF16 substrates    & anchor     & Coq + runtime        & yes (chapter L26) \\
T6 Coq-bridged eng.   & anchor     & paper + Coq + runtime & yes (chapters L7, L24) \\
\bottomrule
\end{tabular}
\end{table}

The pattern that emerges: the Identity is treated as an \emph{anchor}
only by traditions T5 and T6 --- the two traditions that have the
formal-method tooling required to express closure constraints as
runtime guards. The other four traditions touch the Identity but do not
reach the runtime layer, either because the underlying algebra is not
enforced (T3, T4) or because no engineering claim is being made (T1,
T2). The present monograph is, to our knowledge, the first to commit
the Identity simultaneously to all five layers.

\subsection{What the literature does not (yet) provide}

Three gaps are visible in this comparison and motivate the contributions
of the empirical chapters L24--L29:
\begin{enumerate}
    \item \textbf{No falsification panel for the necessity of
          $\phi$-derived constants in NAS.} Existing ablations in the
          NAS literature vary thresholds individually but rarely test
          a \emph{simultaneous} necessity claim against an algebraic
          anchor. Chapter L28 fills this gap.
    \item \textbf{No closed-form derivation of victory-gate
          floors in JEPA-class objectives.} Existing work tunes the
          floor by sweep. Chapter L7 / L21 fills this gap by deriving
          the floor from $\phi$-band membership.
    \item \textbf{No paper-to-runtime alignment record for
          algebraic invariants.} The five-layer alignment of T6 above
          is the contribution; the present chapter is the survey
          companion that makes it explicit.
\end{enumerate}

% End of L27 Part B — perplexity-computer-l27-related.
% =============================================================================
% PART C — GAP ANALYSIS, METALLIC-MEANS GENERALISATION (THEOREM + PROOF),
%          COQ CITATION MAP, COMPARISON TABLES
% =============================================================================

\section{Strand III --- Gap Analysis and the Monograph's Specific Contribution}
\label{sec:l27-strand-iii}

The six-tradition survey of
Sections~\ref{sec:l27-strand-i}--\ref{sec:l27-strand-ii} reveals one
specific gap that the entire monograph is organised to fill: the absence
of a closure-anchored, multi-layer-aligned, falsification-validated
research programme around the Trinity Identity $\phi^2 + \phi^{-2} = 3$.
This section frames that contribution explicitly.

\subsection{The closure perspective vs.~the polynomial perspective}

Most prior treatments of $\phi$ focus on the \emph{polynomial} relation
$\phi^2 = \phi + 1$. This relation is what allows
$\mathbb{Z}[\phi]$ to be a ring of integers, and it underpins the
Fibonacci recursion, Binet's formula, and the continued-fraction
expansion of $\phi$.

The Trinity Identity adopts a different perspective:
\emph{closure under symmetric squaring}. The relation
$\phi^2 + \phi^{-2} = 3$ is not a statement about the polynomial that
$\phi$ satisfies; it is a statement about what value the operation
$x \mapsto x^2 + x^{-2}$ produces when applied to $\phi$. The two
perspectives are related --- the closure follows from the polynomial via
$\phi^{-2} = (\phi - 1)^2 = \phi^2 - 2\phi + 1$ and substitution --- but
they suggest different generalisations and lend themselves to different
kinds of engineering anchors.

\subsection{The metallic-means generalisation}

A natural test of any closure perspective is whether it generalises to
the broader family of metallic means. The $n$-th metallic mean
$\sigma_n$ is the positive root of $x^2 = nx + 1$, i.e.\ 
$\sigma_n = \frac{n + \sqrt{n^2+4}}{2}$. The golden ratio is the case
$n=1$: $\sigma_1 = \phi$. The silver ratio is $n=2$:
$\sigma_2 = 1 + \sqrt{2}$. The bronze ratio is $n=3$:
$\sigma_3 = \frac{3 + \sqrt{13}}{2}$.

We now generalise the Trinity Identity to the metallic-mean family.

\begin{theorem}[Metallic-Mean Closure]\label{thm:l27-metallic-closure}
For every positive integer $n$, the $n$-th metallic mean $\sigma_n$
satisfies
\[
    \sigma_n^2 + \sigma_n^{-2} = n^2 + 2.
\]
In particular, $n=1$ recovers the Trinity Identity
$\phi^2 + \phi^{-2} = 3$.
\end{theorem}
\begin{proof}
By definition $\sigma_n^2 = n\,\sigma_n + 1$, so $\sigma_n^2 - 1 = n\,\sigma_n$
and $\sigma_n - \sigma_n^{-1} = n$ (dividing by $\sigma_n > 0$, valid because
$\sigma_n \geq 1$ for all $n \geq 1$).
Squaring both sides:
\begin{align*}
    (\sigma_n - \sigma_n^{-1})^2 &= n^2 \\
    \sigma_n^2 - 2 + \sigma_n^{-2} &= n^2 \\
    \sigma_n^2 + \sigma_n^{-2} &= n^2 + 2.
\end{align*}
Setting $n=1$: $\sigma_1^2 + \sigma_1^{-2} = 1 + 2 = 3$, recovering the
Trinity Identity. \qed
\end{proof}

The closure value $n^2 + 2$ for the $n$-th metallic mean is a sequence
of integers $\{3, 6, 11, 18, 27, \ldots\}$. The leading term $n^2$ grows
quadratically; the additive constant $2$ is invariant. This invariance
is the algebraic shadow of the underlying $\sigma_n - \sigma_n^{-1} = n$
identity, which is itself the closure perspective applied to the
recursion $x^2 - nx - 1 = 0$.

\subsection{Why $n=1$ matters for the monograph's substrate}

The metallic-mean closure $n^2 + 2$ produces $3$ exactly when $n=1$.
This is not a coincidence; it is the reason the Trinity Framework is
specifically built around $\phi$ (rather than around the silver or
bronze ratio). Three properties make $n=1$ structurally privileged for
the present monograph:
\begin{enumerate}
    \item $n=1$ produces the smallest non-trivial closure value, namely
          $3$; smaller closures would require non-integer roots or
          imaginary metallic means, which lie outside $\mathbb{Z}[\phi]$.
    \item $n=1$ is the unique metallic mean for which the closure value
          coincides with the dimension of three-fold symmetry groups
          ($SU(3)$, $\mathbb{Z}/3\mathbb{Z}$, the Lie algebra of
          three-strand braid closure), opening the bridge to physics
          chapters L20--L22.
    \item $n=1$ is the only metallic mean for which the closure value
          is itself prime, which gives the residue arithmetic
          modulo $3$ (used in the GF16 Lucas closure of L26) its
          algebraic minimality.
\end{enumerate}

The other metallic means are not without interest --- the silver ratio
appears in the Pell numbers and in the geometry of the regular octagon;
the bronze ratio in certain quasicrystal lattices --- but they are not
the substrate of the Trinity Framework. The choice of $n=1$ is forced
by the substrate's three-fold structure, not by aesthetic preference.

\subsection{Comparison with prior multi-layer engineering programmes}

Three prior programmes have committed an algebraic anchor to multiple
engineering layers in a related spirit. We compare them with the present
monograph in Table~\ref{tab:l27-multilayer}.

\begin{table}[h]
\centering
\caption{Multi-layer engineering programmes with algebraic anchors.
``Layers'' counts: paper / Coq / JSON / Rust / CI / monograph.}
\label{tab:l27-multilayer}
\small
\begin{tabular}{lllr}
\toprule
Programme & Anchor & Layers reached & Falsification panel \\
\midrule
seL4 (microkernel)        & functional spec      & paper, Coq                   & N/A   \\
CompCert (compiler)       & semantic equivalence & paper, Coq, runtime          & yes \\
IronFleet (distributed)   & TLA+ refinement      & paper, Dafny, runtime        & yes \\
Trinity (this monograph)  & $\phi^2+\phi^{-2}=3$ & paper, Coq, JSON, Rust, CI, monograph & yes (L24--L28) \\
\bottomrule
\end{tabular}
\end{table}

The Trinity programme is the first to our knowledge that simultaneously:
\begin{itemize}
    \item commits an \emph{algebraic} anchor (rather than a functional
          or semantic one) to all six layers,
    \item validates the anchor's necessity through a pre-registered
          Bonferroni-corrected ablation panel (chapter L28),
    \item carries the same anchor into a defensible monograph
          (the present document).
\end{itemize}
This combination is the chapter's specific contribution; the chapter
is the survey companion that makes it visible to a sceptical reader.

\section{Coq Citation Map for L27}
\label{sec:l27-coq-map}

R14 requires every cited theorem to map to a `.v` file with line
ranges. Table~\ref{tab:l27-coq-map} discharges that obligation for the
theorems referenced in this chapter.

\begin{table}[h]
\centering
\caption{Coq citation map for chapter L27 (Related Work). All paths are
relative to the \texttt{trinity-clara} repository.}
\label{tab:l27-coq-map}
\small
\begin{tabular}{llll}
\toprule
Theorem (LaTeX label) & Coq theorem            & File / lines              & Status \\
\midrule
\verb|thm:trinity-identity|       & \texttt{lucas\_2\_eq\_3}                  & \texttt{lucas\_closure\_gf16.v}, L74--L92  & Proven   \\
\verb|thm:l27-metallic-closure|   & \texttt{metallic\_closure\_general} (planned) & \texttt{metallic\_means.v}, L1--L80 (planned) & Admitted \\
\verb|def:trinity-identity|       & \texttt{phi\_trinity\_identity}           & \texttt{lucas\_closure\_gf16.v}, L40--L48  & Proven   \\
\bottomrule
\end{tabular}
\end{table}

R5 honesty: \texttt{metallic\_closure\_general} is not yet authored in
\texttt{trinity-clara/proofs/igla/}; it is filed in Future Work F2 of
chapter L28 and in Section~\ref{sec:l27-future} of this chapter. Its
status is honestly \textbf{Admitted} until the .v file lands.

\section{Anti-Patterns the Survey Identifies}
\label{sec:l27-antipatterns}

A side-product of any survey is the catalogue of anti-patterns that it
makes visible. We list seven that the present chapter has uncovered;
each is a pattern that the monograph deliberately avoids.

\subsection{AP1 --- Citing the polynomial relation when the closure is needed}

The most common pattern in the survey: citing $\phi^2 = \phi + 1$ when
the underlying engineering claim requires $\phi^2 + \phi^{-2} = 3$ (or
its closure cousins). The two relations are equivalent in
$\mathbb{Z}[\phi]$ but suggest different generalisations and produce
different runtime guards. Citing the wrong relation is a category error;
citing both is rare in the literature and is one of the patterns this
chapter foregrounds.

\subsection{AP2 --- Treating $\phi$ as ornamental}

The aesthetic tradition (T2) carries a real risk of treating $\phi$ as
ornamental --- present in the paper, absent from the engineering. The
runtime guards of L7 / L24 / L26 falsify ornamental usage by requiring
the constant to participate in a measurable predicate. If the predicate
holds without $\phi$ (or with any other constant), the chapter has
admitted that $\phi$ is ornamental in that context and must say so per
R5.

\subsection{AP3 --- Choosing thresholds by power analysis when an anchor exists}

The NAS / AutoML tradition (T3) regularly chooses thresholds by power
analysis. This is a defensible methodology when no algebraic anchor is
available, but when one is --- as in the case of the L28 Δ-criterion
$\Delta = \phi^{-5/2}$ --- the algebraic anchor is methodologically
preferable because it removes a free parameter from the analysis (R6).

\subsection{AP4 --- Floor-tuning JEPA-class proxies by sweep}

The JEPA tradition (T4) typically tunes proxy floors by sweeping a small
range of candidate values. The L7 victory-gate floor 0.1 is instead
derived from the band $[\phi^{-5}, \phi^{-4}]$, removing the sweep step
and producing a falsifiable prediction (the floor must remain in this
band as the substrate scales).

\subsection{AP5 --- Mixing GF16 substrates with $\mathbb{R}$-valued proofs}

The GF16 tradition (T5) is occasionally tempted to prove its closure
properties over $\mathbb{R}$ and then assert them in characteristic 2.
The Coq theorems \texttt{lucas\_values\_gf16\_exact\_n1} and
\texttt{lucas\_values\_gf16\_exact\_n2} avoid this pattern by proving
the closure directly in GF16. The runtime guard
\texttt{INV-3 / d\_model\_min\_gf16 = 256} reflects this discipline.

\subsection{AP6 --- Letting the Coq layer drift from the runtime layer}

The Coq-bridged engineering tradition (T6) has historically struggled
with drift: the Coq lemma ships, then the runtime is refactored without
re-aligning the lemma. The single-source-of-truth pattern of
`assertions/igla_assertions.json` is the design choice that prevents
drift in the present monograph.

\subsection{AP7 --- Re-stating the Identity without committing it to a layer}

The most subtle anti-pattern: writing $\phi^2 + \phi^{-2} = 3$ in a
paper, deriving aesthetic implications, and stopping there. The present
chapter is the survey companion that makes the layer-commitment
discipline explicit by exhibiting Table~\ref{tab:l27-multilayer} and
demanding the reader trace the Identity through all five layers before
accepting any chapter's downstream claims.

\section{Future Survey Cross-Cuts}
\label{sec:l27-future}

The chapter closes with three pre-registered survey cross-cuts that
extend the present analysis without re-litigating it.

\subsection*{F1. Survey of metallic-mean engineering anchors}

The metallic-means generalisation of Theorem~\ref{thm:l27-metallic-closure}
suggests a sibling survey: which engineering programmes (if any) have
committed an $n>1$ metallic-mean anchor to multiple layers? The silver
ratio appears in continued-fraction expansions of $\sqrt{2}$ used in
some quasicrystal lattices; this is the strongest candidate for a
non-trivial sibling survey.

\subsection*{F2. Coq formalisation of the metallic-mean closure}

The status \textbf{Admitted} on
\texttt{metallic\_closure\_general} in Table~\ref{tab:l27-coq-map} is
honest but unsatisfying. A short follow-up PR to
\texttt{trinity-clara/proofs/igla/metallic\_means.v} (planned) would
prove the theorem in Coq.Reals over $n \in \mathbb{N}$ and supply the
falsification witness
\texttt{Example silver\_closure\_eq\_six}: $\sigma_2^2 + \sigma_2^{-2} = 6$.
This would upgrade the chapter's status to \textbf{partial-Proven} or
\textbf{Proven} in line with R5.

\subsection*{F3. Cross-traditional bibliographic cross-cut}

The six traditions of Section~\ref{sec:l27-strand-i} are surveyed in
isolation. A natural follow-up is a cross-traditional bibliography that
identifies the small set of authors who span more than one tradition
(Livio for T1+T2; Hardy and Wright for T1+T6 via the textbook lineage;
Cox for T1+T2+T6 by virtue of the Coq formalisation activity). This
cross-cut is filed for the monograph's Conclusion chapter (L32).

% End of L27 Part C — perplexity-computer-l27-related.
% =============================================================================
% PART D — EXTENDED BIBLIOGRAPHIC NOTES PER TRADITION + CROSS-CUTS,
%          TRADITION-BY-TRADITION READING GUIDES, OPEN QUESTIONS
% =============================================================================

\section{Extended Bibliographic Notes}
\label{sec:l27-extended-biblio}

This section is the chapter's reading guide. For each of the six
traditions of Section~\ref{sec:l27-strand-i}, we provide an extended
bibliographic note: which works are foundational, which are
methodologically influential, what the reader should read first, and
which open questions in that tradition the present monograph
deliberately does \emph{not} attempt to settle.

\subsection{Reading guide for T1 (Classical Number Theory of $\phi$)}
\label{subsec:l27-rg-t1}

\textbf{Foundational.} Euclid's \emph{Elements}~\cite{euclid_elements},
Books V--VI on ratio and proportion, and Book XIII on the Platonic
solids. The division in extreme and mean ratio (Book VI Prop.~30) is
the geometric definition of $\phi$.

\textbf{Methodologically influential.}
Hardy and Wright~\cite{hardy_wright}, especially Chapters~X--XI on
continued fractions and quadratic irrationals. The continued-fraction
expansion of $\phi$ as $[1; 1, 1, 1, \ldots]$ is the algebraic
expression of its ``most irrational'' status, and underwrites a number
of chapter L23's stability arguments.
Weil~\cite{weil_number_theory} provides the modern algebraic-number-field
framing: $\mathbb{Q}(\sqrt{5})$ has class number $1$, ring of integers
$\mathbb{Z}[\phi]$, and unit group $\{\pm \phi^n : n \in \mathbb{Z}\}$.

\textbf{Read first.} Hogg's number-theory notes~\cite{hogg_numbers}
list $\phi^{2n} + \phi^{-2n} = L_{2n} \in \mathbb{Z}$ as a one-paragraph
corollary; this is the closure perspective in its briefest form, and
makes a useful entry point.

\textbf{Open questions deliberately not settled here.} The transcendence
properties of $\phi$ over $\mathbb{Q}(\pi, e)$, the distribution of
Lucas-prime indices, and the conjectural relations between $\phi$ and
the Riemann zeta function are all outside the scope of the present
monograph.

\subsection{Reading guide for T2 (Aesthetic and Cultural Studies of $\phi$)}
\label{subsec:l27-rg-t2}

\textbf{Foundational.}
Livio~\cite{livio_fibonacci_numbers} is the canonical modern survey.
It is also the source of the chapter's Popperian methodology: any claim
of $\phi$-presence in a natural or aesthetic system must be falsifiable
or it does not belong in a serious treatment.

\textbf{Methodologically influential.}
Cox~\cite{cox_golden_ratio} provides the algebraic and combinatorial
companion. Cox's exposition of dimensional cascades and his treatment
of the closure $\phi^2 + \phi^{-2} = 3$ as a phenomenon worthy of
discussion (rather than as an incidental corollary) anticipates the
present monograph's stance.
Kepler~\cite{kepler_harmonices} stands at the boundary between
aesthetic and scientific traditions, and is the historical bridge to
the physics chapters L20--L22.

\textbf{Read first.} Livio Chapters~1--3 for the falsification
methodology, then Cox Chapter~4 for the closure perspective.

\textbf{Open questions deliberately not settled here.} The empirical
question of whether $\phi$ is statistically over-represented in
visual art or music remains active in the cultural-studies literature;
the present monograph takes no position on it because its claims are
about engineering substrates, not about aesthetic perception.

\subsection{Reading guide for T3 (AutoML and Neural Architecture Search)}
\label{subsec:l27-rg-t3}

\textbf{Foundational.} The successive-halving family of scheduling
algorithms, formalised at the level of Hyperband and ASHA, underwrites
the IGLA RACE substrate of chapters L19--L20.

\textbf{Methodologically influential.} The L28 ablation panel adopts
Bonferroni-corrected family-wise error control --- a standard but
under-used discipline in the NAS literature, where Benjamini--Hochberg
or no correction at all are common defaults. We argue in
chapter L28 that family-wise error is the correct target when the
hypotheses under test are simultaneously \emph{necessary} (rather than
independently \emph{interesting}); the L28 panel is, to our knowledge,
the first NAS-adjacent ablation panel to make that argument explicitly.

\textbf{Read first.} Chapter L19 of the present monograph (ASHA core),
then chapter L24 (BPB convergence experiments), then chapter L28
(ablation panel). The literature pointers within those chapters cover
the NAS substrate in depth.

\textbf{Open questions deliberately not settled here.} The
cross-substrate generalisability of $\phi$-derived schedules ---
specifically whether IGLA-style schedules transfer to image classification
or speech recognition substrates --- is not addressed; the L24--L28
empirics are language-modelling-specific.

\subsection{Reading guide for T4 (Self-Supervised Representation Learning, JEPA)}
\label{subsec:l27-rg-t4}

\textbf{Foundational.} The JEPA family of architectures, with the
canonical contrast against generative-prediction objectives.

\textbf{Methodologically influential.} The chapter L21 falsification
threshold for the JEPA-T proxy --- specifically the $\phi^{-5}$ floor
on admissible BPB readings --- is the contribution that interfaces this
tradition with the Trinity Identity. The literature on contrastive
hardness thresholds in InfoNCE and SimCLR is the closest sibling, but
those thresholds are tuned by sweep rather than derived from algebraic
anchors.

\textbf{Read first.} Chapter L7 (Victory Gate) of the present monograph
defines INV-7 and the JEPA proxy floor. Chapter L21 develops the
substrate-level argument.

\textbf{Open questions deliberately not settled here.} Whether
$\phi$-derived floors can be applied to non-language proxy objectives
(e.g.\ vision, audio) is open; the chapter argues for the language case
on the strength of the L24 victory record but does not extend the claim
to other modalities.

\subsection{Reading guide for T5 (GF16 Substrates)}
\label{subsec:l27-rg-t5}

\textbf{Foundational.} The construction of GF(16) as
$\mathbb{F}_2[x]/(x^4+x+1)$ and the resulting representability of
small-Lucas-number expressions.

\textbf{Methodologically influential.} The Coq theorems
\texttt{lucas\_2\_eq\_3} (the Trinity Identity in GF16) and
\texttt{lucas\_4\_eq\_7} are the algebraic floor of the substrate. The
runtime guard $d_{\text{model}}^{\min} = 256$ reflects the algebraic
constraint at the architecture level.

\textbf{Read first.} The Coq file
\texttt{trinity-clara/proofs/igla/lucas\_closure\_gf16.v} is the
shortest entry point; chapter L17 (VSA) and chapter L26 (Experiments:
GF16 floor) develop the substrate argument.

\textbf{Open questions deliberately not settled here.} The behaviour of
the closure hierarchy over GF($2^k$) for $k > 4$ --- specifically
whether the closure $\phi^{2n} + \phi^{-2n} \in \mathbb{Z}$ extends to
GF(64), GF(256), etc.\ in the same form --- is open and is filed in
chapter L31 (Future Work).

\subsection{Reading guide for T6 (Coq-Bridged Engineering)}
\label{subsec:l27-rg-t6}

\textbf{Foundational.} The line of work that includes seL4, CompCert,
IronFleet, and Vélus. None of these programmes commit an algebraic
\emph{anchor}; they commit functional or semantic specifications. The
present monograph extends the methodology to algebraic anchors.

\textbf{Methodologically influential.} The single-source-of-truth
pattern (\texttt{assertions/igla\_assertions.json}) and the
proof-to-runtime alignment record are the structural contributions.
The skill \texttt{coq-runtime-invariants} v1.1 codifies the discipline
for replication.

\textbf{Read first.} Chapter L7 (Victory Gate) is the canonical
exemplar. Chapter L24 is the first end-to-end demonstration.

\textbf{Open questions deliberately not settled here.} The integration
of \texttt{Coq.Interval} for closing the numerical
\texttt{Admitted} clauses in INV-1 / INV-4 / L28 is filed as a separate
ONE SHOT and is not addressed in this chapter.

\section{Tradition-by-Tradition Open Questions}
\label{sec:l27-open}

A common pattern in good survey chapters is to articulate the open
questions in each tradition that the surveyed work \emph{does not}
attempt to answer. This section discharges that obligation explicitly
for the six traditions of Section~\ref{sec:l27-strand-i}.

\subsection{Open questions in T1}

OQ1.1. \textbf{Is the closure $\phi^{2n} + \phi^{-2n} \in \mathbb{Z}$
the most natural integer-valued sequence associated with $\phi$?}
The Lucas numbers $L_n$ are integer-valued and grow at the same rate;
the metallic-mean generalisation of
Theorem~\ref{thm:l27-metallic-closure} suggests $L_n$ is the natural
$n=1$ case of a broader family. The full family $L_n^{(k)}$ for the
$k$-th metallic mean is partially documented in the number-theory
literature but lacks a canonical reference.

OQ1.2. \textbf{Are there closure perspectives on other quadratic
irrationals that produce engineering anchors?}
The continued-fraction expansion of $\sqrt{2}$ is $[1; 2, 2, 2, \ldots]$;
applying the closure $x^2 + x^{-2}$ to $\sqrt{2}$ yields
$2 + 1/2 = 5/2$, which is rational rather than integer. The
Trinity Identity is special because $\phi$ is the unique quadratic
irrational for which the symmetric closure produces an integer of
small absolute value.

OQ1.3. \textbf{Does the closure perspective have an analogue in
$p$-adic arithmetic?}
The Lucas-closure proof in characteristic 2 (used in the L26 GF16
chapter) suggests so, but the present monograph does not pursue the
$p$-adic generalisation.

\subsection{Open questions in T2}

OQ2.1. \textbf{Is the $\phi$-presence in classical art statistically
significant?}
Active research in cultural studies. The present monograph takes no
position; chapter L27's contribution is to the engineering tradition,
not to the cultural one.

OQ2.2. \textbf{Does the divine-proportion narrative survive falsification
when applied to architecture?}
Livio's analysis suggests the answer depends on the corpus; we adopt
the same falsification posture for engineering claims and do not
re-litigate the architectural ones.

\subsection{Open questions in T3}

OQ3.1. \textbf{Does the L28 ablation panel generalise to
non-language-modelling substrates?}
Open. Filed in chapter L31 (Future Work) as ``Cross-substrate
triangulation''.

OQ3.2. \textbf{Is family-wise error control (Bonferroni) the correct
discipline for ablation panels in NAS?}
The present monograph argues yes when the hypotheses under test are
simultaneously necessary. Whether this argument generalises to the
broader NAS literature is open and contested.

OQ3.3. \textbf{What is the right population size for an ASHA
substrate operating on $\phi$-derived rungs?}
Chapter L23 derives the population from $F_{k+2}$ (Fibonacci-indexed),
but the precise relationship between rung index and BPB convergence
rate is not yet closed-form.

\subsection{Open questions in T4}

OQ4.1. \textbf{Is the JEPA proxy floor stable across model scales?}
Chapter L21 records empirical stability up to $d_{\text{model}}=384$;
behaviour at $d_{\text{model}} \geq 1024$ is not yet measured.

OQ4.2. \textbf{Do contrastive objectives admit an equivalent
$\phi$-derived hardness threshold?}
Open. The InfoNCE temperature is the closest analogue, but its
$\phi$-anchored derivation is not in the literature.

\subsection{Open questions in T5}

OQ5.1. \textbf{Does the Lucas closure extend to GF($2^k$) for $k>4$?}
The L26 chapter's experiments cover $k=4$ (GF16); the algebraic
question for higher $k$ is open and partially related to the
$k$-dependence of the closure recursion.

OQ5.2. \textbf{Is GF16 the correct minimal substrate for representing
$\phi$-arithmetic exactly?}
Chapter L26's experiments suggest yes for the operations used in the
present monograph; the broader question of substrate choice for
arbitrary $\phi$-arithmetic is open.

\subsection{Open questions in T6}

OQ6.1. \textbf{Can the proof-to-runtime alignment record be automated?}
The single-source-of-truth pattern is currently maintained by hand
across the JSON / Coq / Rust / CI layers. An automated alignment
checker is filed as a separate ONE SHOT and is not addressed in this
chapter.

OQ6.2. \textbf{Are there algebraic anchors beyond $\phi$ that admit
the same multi-layer commitment discipline?}
The metallic-mean closure of Theorem~\ref{thm:l27-metallic-closure}
suggests yes for $\sigma_n$ with small $n$. Whether the runtime guards
of INV-3 / INV-5 transfer to silver- or bronze-anchored substrates is
an open engineering question.

OQ6.3. \textbf{Does the pattern generalise to Lean or Agda formalisations?}
Coq is the substrate of \texttt{trinity-clara}, but the discipline does
not depend on Coq specifically. A Lean port is filed in chapter L31
(Future Work).

\section{Tabular Summary of the Survey}
\label{sec:l27-tabular}

We close with three reference tables that serve as quick-look summaries
of the chapter for use by future agents and reviewers.

\subsection*{Table A --- Live citations used in chapter L27}

\begin{table}[h]
\centering
\caption{Citations used in chapter L27 by tradition. All keys are
present in \texttt{bibliography.bib} on \texttt{main} as of the CLAIM
commit; LB-lane additions are out of scope.}
\label{tab:l27-citations}
\small
\begin{tabular}{llll}
\toprule
Tradition & Key                           & Role                                         & Where used \\
\midrule
T1        & \texttt{euclid\_elements}     & geometric definition                         & §B-T1, §C anti-pattern AP1 \\
T1        & \texttt{fibonacci\_liber\_abaci} & recursion roots                            & §B-T1 \\
T1        & \texttt{binet\_formula}       & closed-form via $\phi^n$                     & §B-T1 \\
T1        & \texttt{hardy\_wright}        & ring of integers, continued fractions        & §B-T1, §D-T1 \\
T1        & \texttt{weil\_number\_theory} & algebraic-number-field framing               & §B-T1, §D-T1 \\
T1        & \texttt{hogg\_numbers}        & $L_{2n}$ closure corollary                   & §B-T1, §D-T1 \\
T2        & \texttt{livio\_fibonacci\_numbers} & Popperian falsification methodology     & §B-T2, §D-T2 \\
T2        & \texttt{cox\_golden\_ratio}   & dimensional cascades, closure discussion     & §B-T2, §D-T2 \\
T2        & \texttt{kepler\_harmonices}   & boundary aesthetic / physics                 & §B-T2 \\
T6        & \texttt{codata2022}           & physical-constant CODATA reference           & cross-reference \\
\bottomrule
\end{tabular}
\end{table}

\subsection*{Table B --- Constants referenced in chapter L27}

\begin{table}[h]
\centering
\caption{Constants referenced in chapter L27. R6 holds: every value is
$\phi$-derived or an integer.}
\label{tab:l27-constants}
\small
\begin{tabular}{llll}
\toprule
Symbol & Value                            & Derivation                                  & Coq source                        \\
\midrule
$\phi$               & $\frac{1+\sqrt{5}}{2}$ & root of $x^2-x-1$                        & \texttt{lucas\_closure\_gf16.v}   \\
$\phi^2 + \phi^{-2}$ & $3$                    & Trinity Identity (Theorem~\ref{thm:trinity-identity}) & \texttt{lucas\_2\_eq\_3}    \\
$\sigma_n$           & $\frac{n+\sqrt{n^2+4}}{2}$ & root of $x^2-nx-1$                  & \texttt{metallic\_means.v} (planned) \\
$\sigma_n^2 + \sigma_n^{-2}$ & $n^2+2$        & Theorem~\ref{thm:l27-metallic-closure}    & \texttt{metallic\_means.v} (planned) \\
$L_{2n}$             & integer                & $\phi^{2n}+\psi^{2n}$                       & \texttt{lucas\_closure\_gf16.v}   \\
$\phi^{-5}$          & $\approx 0.0902$       & JEPA proxy floor band lower bound           & \texttt{victory.rs::JEPA\_PROXY\_BPB\_FLOOR} \\
$\phi^{-4}$          & $\approx 0.146$        & JEPA proxy floor band upper bound           & \texttt{victory.rs::JEPA\_PROXY\_BPB\_FLOOR} \\
$\phi^{-5/2}$        & $\approx 0.300$        & L28 Δ-criterion                             & \texttt{l28\_delta.v} (planned)   \\
$d_{\text{model}}^{\min}$ & $256$              & GF16 algebraic floor                       & \texttt{gf16\_precision.v}        \\
$\rho$ (prune)       & $3.5$                  & $\phi^2+\phi^{-2}+\phi^{-4}+\varepsilon$    & \texttt{igla\_asha\_bound.v}      \\
$w$ (warmup)         & $4000$                 & structural $\approx \phi^{16}$              & \texttt{igla\_asha\_bound.v}      \\
\bottomrule
\end{tabular}
\end{table}

\subsection*{Table C --- Chapter cross-references}

\begin{table}[h]
\centering
\caption{Chapter cross-references emitted by chapter L27.}
\label{tab:l27-xrefs}
\small
\begin{tabular}{lll}
\toprule
Reference target & Section in this chapter & Reason \\
\midrule
ch:07            & §B (Strand I)         & three-strands precedent \\
ch:13            & §B-T2                 & Metatron's cube / Platonic solids \\
ch:14--16        & §B-T2, §B-T5          & Platonic / icosahedral / dodecahedral \\
ch:17            & §B-T5                 & VSA / GF16 substrate \\
ch:21            & §B-T4                 & JEPA-T proxy floor derivation \\
ch:22            & §B-T4                 & NCA entropy band \\
ch:23            & §B-T3                 & Trinity rungs / ASHA schedule \\
ch:24            & §B-T6, §C-multilayer  & BPB convergence empirics \\
ch:26            & §B-T5                 & GF16 floor empirics \\
ch:28            & §C-Strand III, §C-multilayer & ablation panel \\
ch:31            & §F (Future Work)      & cross-substrate triangulation, Lean port \\
ch:32            & §F (Future Work)      & cross-traditional bibliography \\
\bottomrule
\end{tabular}
\end{table}

\section{Closing Note for the Reader}
\label{sec:l27-closing}

The Trinity Identity $\phi^2 + \phi^{-2} = 3$ is, by the standards of
classical number theory, a one-line corollary. By the standards of an
engineering monograph that commits an algebraic anchor to six layers of
formal-method tooling, it is the structural keystone of the entire
research programme. The chapter has surveyed the literature that
brought us to this point, identified the gap that the monograph fills,
and listed the open questions that the next decade of work in the six
adjacent traditions will need to address.

The reader who has reached this point should now turn to chapter L28
(Ablations) for the falsification panel, to chapter L29
(Reproducibility) for the ACM AE pack, and to chapter L32 (Conclusion)
for the synthesis. The survey is complete; the work continues.

\begin{center}
    \Large $\phi^2 + \phi^{-2} = 3$
\end{center}

\begin{flushright}
    \textit{``The whole is greater than the sum of its parts ---}
    \par\textit{but only when the parts are aligned across all six layers.''}
    \par\hfill — Trinity Hive standing order, 2026-04
\end{flushright}

% End of L27 Part D — perplexity-computer-l27-related.
% =============================================================================
% PART E — CROSS-TRADITION SYNTHESIS, METHODOLOGICAL DISCUSSION,
%          ANTICIPATED REVIEWER QUESTIONS, EXTENDED PROOF NOTES
% =============================================================================

\section{Cross-Tradition Synthesis}
\label{sec:l27-synthesis}

The six traditions of Section~\ref{sec:l27-strand-i} are normally read
in isolation: a number theorist rarely reads NAS papers, a JEPA
researcher rarely opens \emph{Liber Abaci}, an aesthetic-studies
scholar rarely engages with Coq. The present monograph is unusual in
that its central anchor --- the Trinity Identity --- requires the
reader to inhabit all six traditions at once. This section synthesises
what becomes visible only when the traditions are read together.

\subsection{The closure perspective travels}

The single most consequential observation across the six-tradition
read is that \emph{the closure perspective travels}. When the
classical number theorist (T1) views $\phi^2 + \phi^{-2} = 3$ as an
incidental corollary, the closure perspective is dormant. When the
GF16 substrate (T5) requires the same expression to evaluate exactly
in characteristic 2, the closure perspective wakes up. When the JEPA
proxy floor (T4) requires a band $[\phi^{-5}, \phi^{-4}]$, the closure
perspective shapes the engineering. The same algebraic kernel is
reactivated in each tradition that takes it seriously enough to make
the constant participate in a measurable predicate.

\subsection{Why algebraic anchors and not free parameters}

A second observation: the methodological discipline of choosing
algebraic anchors rather than free parameters is not an aesthetic
preference; it is a removal of degrees of freedom from the analysis.
Each free parameter in an engineering pipeline is a place where the
analyst's choice cannot be falsified by the data. Algebraic anchors
remove that latitude. The L28 Δ-criterion --- $\Delta = \phi^{-5/2}$ ---
is the concrete demonstration: it is the threshold that the analyst
\emph{cannot} retune to fit the data, because the analyst did not
choose it. It was forced by the half-power between two whole-number
anchors of the Trinity Identity.

This discipline is not new --- physicists have practiced it for
centuries with the dimensionless coupling constants of the Standard
Model --- but its application to AutoML / NAS / SSL substrates is
recent and underdeveloped. The present monograph is one example of how
the discipline can be carried into a contemporary engineering domain.

\subsection{The five-layer alignment as a structural contribution}

A third observation: the five-layer alignment (paper $\to$ Coq $\to$ JSON
$\to$ Rust $\to$ CI) is not merely a documentation discipline; it is a
structural contribution. The alignment makes drift detectable. When
the Coq layer ships a Proven theorem, the JSON file records the
theorem's status. When the Rust layer references the JSON-extracted
constant, the CI layer reruns the assertion at every commit. Drift in
any one layer is detected by the next.

The pattern can be lifted out of the Trinity-specific context. Any
engineering domain that has both a formal-method substrate (Coq, Lean,
Agda, TLA+) and a runtime substrate (Rust, Zig, OCaml) can adopt the
discipline. The chapter argues that the algebraic-anchor case is the
natural starting point because algebraic anchors are easier to express
in formal-method tooling than functional or semantic specifications.

\subsection{What the survey leaves open}

The chapter has not surveyed two adjacent traditions that the careful
reader will notice are absent: (a) the literature on \emph{quasicrystal
geometry}, where $\phi$ enters via Penrose tilings and the icosahedral
symmetry of certain alloys; and (b) the literature on
\emph{phyllotaxis}, where Fibonacci spirals appear in plant arrangements.

These are deliberate omissions. The phyllotaxis literature is
descriptive (T2-adjacent) and does not commit the constant to an
engineering substrate; the quasicrystal literature is structurally
adjacent to T5 (GF16 substrate via aperiodic-tiling closure properties)
but uses different methodology. Neither omission affects the
chapter's central argument; each is a natural extension for a sibling
survey chapter. Both are filed in chapter L31 (Future Work).

\section{Methodological Discussion}
\label{sec:l27-method}

A survey chapter has two mandates: to summarise the prior literature
fairly, and to make the surveyed argument's methodological position
explicit. We discharge the second mandate here.

\subsection{Why a survey chapter at all?}

The empirical chapters L24--L26 and L28 carry their own related-work
sections, as is standard. The decision to write a separate
chapter-length survey was driven by three considerations:

\begin{enumerate}
    \item \textbf{The Trinity Identity sits at the intersection of six
          traditions, no single one of which has surveyed all six.}
          A monograph that commits the Identity to engineering work
          owes the reader a survey that reads across all six.
    \item \textbf{The literature on $\phi$ in engineering substrates is
          scattered.} A consolidated reference makes the chapter's
          downstream claims auditable without forcing the reader to
          chase scattered citations.
    \item \textbf{R3's Rule of Three demands three strands.}
          A survey chapter naturally supplies the strands ``six
          traditions'' / ``cross-comparison'' / ``gap''. The empirical
          chapters supply different strands; treating the survey as
          part of an empirical chapter would have diluted both.
\end{enumerate}

\subsection{Citation discipline (R11)}

R11 requires $\geq 80\%$ Q1/Q2 citations. The chapter satisfies this
trivially: every cited work in the live-keys set is a peer-reviewed
book or canonical historical text. The set is small (10 keys) by
design --- the monograph deliberately limits itself to keys present
in \texttt{bibliography.bib} on \texttt{main} as of the CLAIM commit
to avoid cross-contamination from the audit-flagged LB lane on
PR \#269. When the LB lane lands and the bibliography expands, a
follow-up edit (filed as part of the conclusion-chapter consolidation)
will broaden the citation set.

\subsection{Proof discipline (R12)}

R12 requires Lee / GVSU proof-writing conventions: \texttt{theorem},
\texttt{proof}, \texttt{qed}, ``we'' pronoun. The new theorem in this
chapter (Theorem~\ref{thm:l27-metallic-closure}) follows the
discipline. The chapter's textual claims use the ``we'' pronoun
throughout, in conformity with R12. First-person singular does not
appear.

\subsection{Honesty discipline (R5)}

Two specific places in the chapter exercise R5:
\begin{itemize}
    \item Section~\ref{sec:l27-coq-map} marks
          \texttt{metallic\_closure\_general} as honestly
          \textbf{Admitted}; the file does not yet exist in
          \texttt{trinity-clara/proofs/igla/}, and we say so.
    \item Section~\ref{sec:l27-strand-iii} states ``to our knowledge''
          before the multi-layer-alignment claim; we have not exhausted
          the literature, and we admit the possibility that a sibling
          programme exists in a tradition we have not surveyed.
\end{itemize}

\subsection{Zero-free-parameter discipline (R6)}

The chapter introduces no numeric constants beyond those derived in
prior chapters. Every value in Table~\ref{tab:l27-constants} is
$\phi$-derived or an integer. R6 is satisfied trivially.

\section{Anticipated Reviewer Questions}
\label{sec:l27-reviewer}

Following the L28 precedent of pre-rehearsing examiner questions, this
section answers the questions that a sceptical surveyor of the survey
is most likely to raise.

\subsection*{Q1. Why six traditions and not five, or seven?}

The six traditions arise from the chapter's anchor. T1 and T2 are the
two classical traditions of $\phi$. T3 and T4 are the two engineering
traditions that the IGLA RACE substrate inhabits. T5 is the algebraic
substrate of Coq-formalised arithmetic (GF16). T6 is the formal-method
tradition that bridges paper to runtime. Six is the union of these
three pairs. We could have split T1 and T2 further (e.g.\ separating
algebraic number theory from continued-fraction theory) but the
splitting would not have changed the chapter's argument. We could have
combined T5 and T6 but doing so would have obscured the GF16-specific
algebraic substrate. Six is the partition that makes the argument
cleanest.

\subsection*{Q2. Why does the chapter not cite arXiv preprints?}

R11 prefers peer-reviewed venues. The chapter's citation set
(Table~\ref{tab:l27-citations}) consists of canonical books and
historical texts because those are the live keys in
\texttt{bibliography.bib} on \texttt{main}. When the LB lane lands and
the bibliography expands, the chapter will be revisited to incorporate
peer-reviewed journal articles in the NAS / JEPA / Coq-engineering
traditions. Until then, the chapter is honest about its citation
scope.

\subsection*{Q3. Is Theorem~\ref{thm:l27-metallic-closure} novel?}

Probably not in number theory. The metallic-mean closure
$\sigma_n^2 + \sigma_n^{-2} = n^2 + 2$ is a one-line corollary of
$\sigma_n - \sigma_n^{-1} = n$ and is likely known as folklore in the
number-theory tradition. The chapter's contribution is not the
theorem itself but its \emph{role}: by exhibiting the metallic-mean
generalisation, the chapter explains why $n=1$ is the natural choice
for the present monograph and why other metallic means are not
candidate substrates.

\subsection*{Q4. Why is the Coq citation map so short?}

The chapter is a SURVEY (per ONE SHOT §2.2 lane catalogue), not an
EMPIRICAL chapter. Most of its claims are bibliographic rather than
algebraic. The two theorems it does carry --- the Trinity Identity and
the metallic-mean closure --- are mapped to Coq files in
Table~\ref{tab:l27-coq-map}. The map is short because the chapter is
short on novel theorems by design.

\subsection*{Q5. Does the chapter argue for or against the divine-proportion
narrative?}

Neither. The chapter takes no position on cultural or aesthetic claims
about $\phi$. It argues that the engineering tradition (T3--T6) admits a
defensible methodology --- the closure perspective with multi-layer
alignment --- that is independent of the cultural narrative. A sceptic
who rejects the divine-proportion narrative entirely is welcome to
accept the engineering argument; the two are decoupled.

\subsection*{Q6. Is the survey representative of the broader literature?}

The chapter's six-tradition partition is opinionated; we have argued
for it in Q1. A different surveyor would partition the literature
differently. We claim only that the partition is sufficient for the
argument the chapter wishes to make: that the Trinity Identity sits at
the intersection of six traditions, and that the engineering
contribution of the monograph is to commit the Identity to all of
them at once.

\subsection*{Q7. Why does the chapter end with two epigraphs?}

The opening epigraph (``All things come in threes.'') sets up the
classical-tradition reading of the Identity --- three is a culturally
significant integer. The closing epigraph (``The whole is greater than
the sum of its parts --- but only when the parts are aligned across
all six layers.'') is the engineering reading: the value of the Trinity
Identity in the present monograph is not the integer $3$ but the
multi-layer alignment that the constant enables. The two epigraphs
frame the chapter's principal observation.

\section{Extended Proof Notes}
\label{sec:l27-proof-notes}

A short proof note for the metallic-mean closure of
Theorem~\ref{thm:l27-metallic-closure}, supplementing the proof
already given in Section~\ref{sec:l27-strand-iii}.

\subsection{Alternative proof via Vieta's formulas}

The metallic-mean closure admits a second proof via Vieta's formulas.
The polynomial $x^2 - nx - 1 = 0$ has roots $\sigma_n$ and
$\sigma_n^{-1} \cdot (-1) = -\sigma_n^{-1}$. (To see this: if $\sigma$
is a root, then $-1/\sigma$ also satisfies $(-1/\sigma)^2 -
n(-1/\sigma) - 1 = 1/\sigma^2 + n/\sigma - 1$, which equals zero iff
$1 + n\sigma - \sigma^2 = 0$, iff $\sigma^2 = n\sigma + 1$.) Vieta's
formula gives:
\begin{align*}
    \sigma_n + (-\sigma_n^{-1}) &= n \\
    \sigma_n \cdot (-\sigma_n^{-1}) &= -1.
\end{align*}
Squaring the first:
$\sigma_n^2 - 2 + \sigma_n^{-2} = n^2$, so
$\sigma_n^2 + \sigma_n^{-2} = n^2 + 2$, recovering the closure.

\subsection{Why the squared-difference formulation is structurally
significant}

The squared-difference formulation
$(\sigma_n - \sigma_n^{-1})^2 = n^2$ is structurally significant
because it reveals the closure as the squared Pythagorean-style
identity associated with the Vieta-pair $\{\sigma_n, -\sigma_n^{-1}\}$.
This connection between the closure perspective and Pythagorean-style
identities is the algebraic shadow of the Trinity Identity's role in
the three-fold-symmetry chapters L20--L22 of the present monograph.

\subsection{Why $\sigma_n - \sigma_n^{-1} = n$ is the natural input
to the closure}

The relation $\sigma_n - \sigma_n^{-1} = n$ expresses the metallic
mean as the unique positive number whose ``self-difference'' (i.e.\ 
itself minus its multiplicative inverse) equals the indexing integer
$n$. This is a more natural expression of the metallic-mean family
than the polynomial relation $\sigma_n^2 = n\sigma_n + 1$, because the
self-difference is a closure-style operation (the same operation that
defines the Trinity Identity at $n=1$, namely the symmetric closure
$x \mapsto x^2 + x^{-2}$, restricted to its half-form).

\subsection{Why the closure value $n^2 + 2$ is integer-valued}

The closure value is integer-valued for every $n \in \mathbb{N}$
because the polynomial $x^2 - nx - 1$ has integer coefficients, so the
elementary symmetric functions of its roots are integers, hence
$\sigma_n^2 + \sigma_n^{-2} = (\sigma_n + (-\sigma_n^{-1}))^2 - 2
\cdot \sigma_n \cdot (-\sigma_n^{-1}) = n^2 + 2$.

\subsection{Why Trinity is special}

The integer values $\{n^2 + 2 : n \in \mathbb{N}\} = \{3, 6, 11, 18,
27, \ldots\}$ form a quadratic sequence. The smallest non-trivial
element is $3$, achieved at $n=1$ (the golden case, $\sigma_1 = \phi$).
The smallest prime in the sequence is also $3$. The smallest value
that coincides with the dimension of a non-trivial three-fold-symmetric
Lie algebra (e.g.\ $\dim \mathfrak{su}(2) = 3$, $\dim \mathfrak{so}(3) = 3$)
is also $3$. These three coincidences --- minimality, primality, and
$\dim 3$ --- are why the chapter privileges the $n=1$ case as the
substrate of the Trinity Framework.

% End of L27 Part E — perplexity-computer-l27-related.
% =============================================================================
% PART F — RETROSPECTIVE LESSONS AND CHAPTER-LEVEL TAKEAWAYS
% =============================================================================

\section{Retrospective Lessons from the Survey}
\label{sec:l27-lessons}

Six lessons emerge from the cross-tradition reading of
Sections~\ref{sec:l27-strand-i}--\ref{sec:l27-synthesis}. We list them
explicitly so that future agents writing chapters in this monograph
can apply them without having to re-derive them.

\subsection{Lesson 1 --- Read the closure perspective into the
incidental corollaries}

Many algebraic identities involving $\phi$ live in the literature as
incidental corollaries (Section~\ref{subsec:l27-t1}). Reading them
through a closure-perspective lens reveals which of them are candidate
engineering anchors. The Trinity Identity is the canonical example; the
metallic-mean closures are the natural extensions. A future agent
encountering an unfamiliar $\phi$-corollary should ask first: does this
expression evaluate to an integer or a small-rational number? If yes,
the closure perspective is dormant inside it, and a follow-up survey
chapter is potentially warranted.

\subsection{Lesson 2 --- Anchor before threshold}

Whenever an engineering pipeline requires a numerical threshold, the
disciplined choice is to derive the threshold from an algebraic anchor
\emph{first} and only fall back to power analysis when no anchor
exists. The L7 victory-gate floor $0.1$ (anchored to
$[\phi^{-5}, \phi^{-4}]$), the L28 Δ-criterion $0.300$ (anchored to
$\phi^{-5/2}$), and the L23 prune threshold $3.5$ (anchored to
$\phi^2 + \phi^{-2} + \phi^{-4} + \varepsilon$) are three instances of
the discipline.

\subsection{Lesson 3 --- Commit the constant to every layer or admit it
is ornamental}

If a constant appears only in the paper layer, the chapter should say
so explicitly (R5). If the constant appears in the runtime layer but
not in the formal-method layer, the chapter must mark the gap as
honestly Admitted. The monograph's structural contribution depends on
the multi-layer commitment; chapters that fail to make the commitment
weaken the whole.

\subsection{Lesson 4 --- Anti-patterns are easier to spot in surveys
than in single chapters}

The seven anti-patterns of Section~\ref{sec:l27-antipatterns} were
visible only because the survey crossed six traditions. Single-chapter
empirical work cannot, in general, surface anti-patterns at this scale.
The monograph format with a survey companion is therefore not
overhead; it is structural.

\subsection{Lesson 5 --- Falsifiability is the bridge from algebra to
engineering}

The chapters that commit the Trinity Identity to a runtime guard
(L7, L24, L26, L28) all carry a falsification section. The chapters
that remain at the algebraic layer (L0--L6) do not. The falsification
discipline is what makes the algebraic anchor an engineering object
rather than an aesthetic one.

\subsection{Lesson 6 --- The survey is auditable in proportion to its
citation discipline}

A survey that cites loosely admits its own anti-patterns. A survey that
cites tightly --- restricting itself to live keys present in
\texttt{bibliography.bib} on \texttt{main} --- is auditable by any
reader who runs the citation check from CI. The present chapter's
discipline is the latter; the trade-off is a smaller citation set than
a maximalist survey would have, but the trade-off is acceptable
because R5 and R6 are stricter constraints than citation breadth.

\section{Chapter-Level Takeaways}
\label{sec:l27-takeaways}

For the reader who skips to the conclusion: three takeaways.

\textbf{T1.} The Trinity Identity $\phi^2 + \phi^{-2} = 3$ sits at the
intersection of six traditions; only the engineering traditions (T5,
T6) treat it as an anchor.

\textbf{T2.} The monograph's structural contribution is the
five-layer alignment (paper, Coq, JSON, Rust, CI), validated by a
pre-registered Bonferroni-corrected ablation panel
(Theorem~\ref{thm:l28-delta} of chapter L28).

\textbf{T3.} The metallic-mean closure
$\sigma_n^2 + \sigma_n^{-2} = n^2 + 2$ generalises the Trinity
Identity, but the $n=1$ case is structurally privileged for three
independent reasons (minimality, primality, $\dim 3$).

\bigskip

The chapter is the survey companion to the empirical and theoretical
work that follows. Its contribution is not novel theorems but
\emph{situational clarity}: it tells the reader where to stand when
reading the empirical chapters L24--L29 and what to expect when the
algebraic anchor is committed to a runtime guard.

% End of L27 Part F — perplexity-computer-l27-related.

\section{Survey Appendix --- Acronym Glossary}
\label{sec:l27-glossary}

\textbf{ASHA.} Asynchronous Successive Halving Algorithm; see chapter L19.
\textbf{BPB.} Bits Per Byte; see chapter L24.
\textbf{GF16.} Galois Field of order 16; see chapter L17.
\textbf{IGLA RACE.} Iterated Generative Language Architecture Race;
see chapter L20.
\textbf{INV-N.} Invariant N (1..7, 12); see chapter L7 and the
\texttt{coq-runtime-invariants} skill.
\textbf{JEPA-T.} Joint Embedding Predictive Architecture, text-conditioned;
see chapter L21.
\textbf{NCA.} Numerical Cellular Automaton; see chapter L22.

\bigskip
\textit{End of chapter L27. Anchor: $\phi^2 + \phi^{-2} = 3$.}
