\chapter{TRI27 DSL and Dual-Target Codegen (Zig + Verilog)}
\label{ch:27b}

\section{Introduction}

The arithmetic core of Trinity S$^3$AI processes weights and activations represented as balanced-ternary vectors. The natural programming substrate for such computations is a three-valued language in which the primitive type \texttt{trit} has exactly three inhabitants: \texttt{Neg} ($-1$), \texttt{Zero} ($0$), and \texttt{Pos} ($+1$). The cardinality of this type is~3---the same integer that appears at the right-hand side of the anchor identity $\varphi^2 + \varphi^{-2} = 3$.

TRI27 (Trinity-27) takes its name from the $27 = 3^3$ possible triples of trit values, the natural unit of computation in the balanced-ternary VM. A TRI27 expression evaluates to a single \texttt{trit} given an environment~$\rho$ mapping variable names to \texttt{trit} values. The two foundational theorems proved for this language---\texttt{eval\_det} (determinism) and \texttt{trit\_exhaustive} (exhaustiveness)---underpin all higher-level correctness properties of the VM and the formal proofs in subsequent chapters.

\section{TRI27 Syntax and Denotational Semantics}

\subsection{Abstract Syntax}

The TRI27 expression language is defined by the following inductive type in Coq:

\begin{verbatim}
Inductive expr : Type :=
  | Lit   : trit -> expr
  | Var   : nat  -> expr
  | Neg3  : expr -> expr
  | Add3  : expr -> expr -> expr
  | Mul3  : expr -> expr -> expr
  | If3   : expr -> expr -> expr -> expr.
\end{verbatim}

The type \texttt{trit} has exactly three constructors: \texttt{Neg}, \texttt{Zero}, \texttt{Pos}.

\subsection{Ternary Arithmetic}

The fundamental ternary operations are defined by the $3 \times 3$ tables implementing $\mathbb{F}_3$ arithmetic. Balanced-ternary addition yields $(a+b) \bmod 3$ with values mapped as $\{-1, 0, +1\}$. The distributive law $a \times_3 (b +_3 c) = (a \times_3 b) +_3 (a \times_3 c)$ holds by inspection and is proved as a derived lemma in \texttt{Trit.v}.

\section{Mechanised Proofs}

\begin{theorem}[Determinism --- \texttt{eval\_det}]
\label{thm:eval-det}
For any expression $e$, environment $\rho$, and trit values $v_1, v_2$: if \texttt{eval e rho = Some v1} and \texttt{eval e rho = Some v2}, then $v_1 = v_2$.
\end{theorem}

\begin{proof}
By structural induction on $e$. The base cases \texttt{Lit t} and \texttt{Var n} return fixed values. For compound expressions, the induction hypothesis provides uniqueness for subexpression results; the ternary operation tables are deterministic (single-valued functions on $\{-1,0,+1\}^2$), so the composite result is unique. Total proof length: 43 lines in \texttt{Semantics.v}.
\end{proof}

\begin{theorem}[Exhaustiveness --- \texttt{trit\_exhaustive}]
\label{thm:trit-exhaustive}
For any $t : \texttt{trit}$, either $t = \texttt{Neg}$ or $t = \texttt{Zero}$ or $t = \texttt{Pos}$.
\end{theorem}

\begin{proof}
By case analysis on the inductive type \texttt{trit}. Since \texttt{trit} has exactly three constructors, \texttt{destruct t} yields three subgoals, each closed by reflexivity. Total proof length: 7 lines in \texttt{Trit.v}.
\end{proof}

\section{Dual-Target Codegen Architecture}

The TRI27 compiler produces code for two primary targets from a single \texttt{.t27} specification:

\begin{table}[h]
\centering
\caption{Codegen targets and metrics}
\label{tab:codegen-targets}
\begin{tabular}{llll}
\toprule
\textbf{Target} & \textbf{Emitter} & \textbf{LoC} & \textbf{Output} \\
\midrule
CPU (Zig)   & \texttt{emit\_zig}    & 600 & \texttt{src/gfN.zig} \\
FPGA (Verilog) & \texttt{emit\_verilog} & 520 & \texttt{fpga/gfN.v} \\
CPU (C)     & \texttt{emit\_c}       & 273 & \texttt{src/gfN.c} \\
CPU (Rust)  & \texttt{emit\_rust}    & --- & \texttt{src/gfN.rs} \\
\bottomrule
\end{tabular}
\end{table}

The compiler itself is implemented in Rust (\texttt{bootstrap/src/compiler.rs}, 20\,814 LoC) with spec-driven modules:
\begin{itemize}
\item \textbf{Lexer}: 597 LoC spec (\texttt{compiler/parser/lexer.t27})
\item \textbf{Parser}: 1\,294 LoC spec (\texttt{compiler/parser/parser.t27})
\item \textbf{AST}: 611 LoC spec (\texttt{compiler/ast.t27})
\item \textbf{Verilog codegen}: 1\,067 LoC spec + 2\,358 LoC FPGA emission spec
\item \textbf{Zig codegen}: 1\,516 LoC spec
\end{itemize}

\subsection{Bit-Identical KPI}

The central correctness property is: from a single \texttt{.t27} spec, both the Zig CPU emitter and the Verilog FPGA emitter produce bit-identical results on $N = 10^5$ random GF16/TF3 operations. The regression suite covers all operations in the GF16 accelerator: multiply, add, MAC, dot product, FFT, IFFT, matmul, and inverse.

\begin{equation}
\forall\, o \in \{\text{GFADD}, \text{GFMUL}, \text{GFMAC}, \ldots\},\, \forall\, (a, b):\; \mathrm{out}_{\mathrm{Zig}}(o, a, b) = \mathrm{out}_{\mathrm{Verilog}}(o, a, b)
\end{equation}

\subsection{Numeric Format Coverage}

The compiler processes 9 numeric format specifications totalling 7\,956 LoC:

\begin{table}[h]
\centering
\caption{Numeric format specs}
\begin{tabular}{lr}
\toprule
\textbf{Spec} & \textbf{LoC} \\
\midrule
\texttt{gf16.t27} (SSOT per L6) & 3\,436 \\
\texttt{tf3.t27}                 & 1\,691 \\
\texttt{gf8.t27}                 & 521 \\
\texttt{gf12.t27}                & 481 \\
\texttt{gf20.t27}                & 468 \\
\texttt{gf24.t27}                & 468 \\
\texttt{gf32.t27}                & 479 \\
\texttt{gf4.t27}                 & 305 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Sealed Spec Registry}

All 33 FPGA module specs carry sealed hashes in \texttt{.trinity/seals/}, with per-target \texttt{gen\_hash\_*} fields for Verilog, Zig, C, and Rust. The GF16 accelerator seal (\texttt{fpga\_Gf16Accel.json}, ring~12, sealed 2026-04-11) records the spec hash, generated code hashes, and test results.

\section{Discussion}

The TRI27 DSL as formalised here is intentionally minimal. The present two theorems establish only determinism and exhaustiveness; a complete verified compiler from TRI27 to FPGA RTL would require additional theorems on type safety, termination, and translation correctness---all planned for v5. The most significant limitation is that the current semantics does not handle variable out-of-scope errors gracefully. A future \texttt{emit\_coq.zig} is planned to generate formal proofs from the same \texttt{.t27} specification, closing the loop between codegen and verification (Ch.~\ref{ch:10}).
