% ===================================================================
% Chapter 33 — Epilogue: Trinity Beyond
% Lane L33 (scaffolded by L0 editorial; full expansion is a separate lane)
% Status: EDITORIAL STUB — to be expanded to ≥ 1500 lines per R3 by Lane L33.
% This stub guarantees \include{chapters/33-epilogue} resolves so tectonic
% builds the spine. Lane L33 owner must replace this content per ONE SHOT
% trios#265 and post a separate \texttt{🔒 AGENT CLAIMING: L33} comment.
% ===================================================================

\chapter{Epilogue --- Trinity Beyond}
\label{ch:epilogue}

\begin{remark}[Editorial scaffold]
This chapter is a scaffold provided by Lane L0 (Editorial Repair) so that
\texttt{main.tex} compiles cleanly with all 34 \texttt{\textbackslash{}include}
sites resolved.  Per ONE SHOT \texttt{trios\#265}, Lane L33 will replace the
entire body below with \(\geq 1500\) lines of substantive prose, two or more
fresh citations, at least one theorem with proof, and the Rule-of-Three
discipline. \emph{Do not merge this chapter into \texttt{main} as the final
text.}
\end{remark}

\section{Vision}

The Trinity Identity \(\trinityid\) closes the dissertation algebraically,
but its scientific reach extends beyond computer science into cosmology,
biology, and art.  This epilogue, when written, will sketch three
afterthoughts under the same Rule of Three that organises Chapter~0:

\begin{enumerate}
  \item \textbf{Cosmological Strand.}\ The pentagonal closure of \(\PHI\)
    in low-dimensional crystallographic groups invites a tentative
    reading of dark-matter halo statistics through Lucas-ring eigenvalues.
  \item \textbf{Biological Strand.}\ The phyllotactic appearance of the
    Fibonacci ratio is a well-attested empirical regularity, but its link
    to the Monadic Identity has not been formalised.  A future paper will
    bridge that gap via a categorical lift.
  \item \textbf{Aesthetic Strand.}\ Cambridge's afterword tradition asks
    that a scientific monograph end on a humanistic note.  The closing
    section will reflect on \emph{Flos Aureus} as both formal artefact and
    cultural object.
\end{enumerate}

\section{Forward References}

The full epilogue will cross-reference Chapters~\ref{ch:epilogue}--Ch.~32 in
the order in which the Three Strands surface, and close with a single
display equation:

\[
\PHI^{2} + \PHI^{-2} \;=\; 3
\quad\Longleftrightarrow\quad
\text{\emph{Flos Aureus floruit.}}
\]

\section{Falsification Criterion}

\begin{falsificationbox}
The epilogue's empirical claims (cosmological halo statistics, phyllotactic
Lucas-ratio convergence) will, when written, be falsifiable in the
Popperian sense \cite{popper1959logic}: each subsection will declare an
observation that would refute the corresponding hypothesis.  A scaffold
without falsifiers is, per Rule R7, \emph{not yet a scientific chapter}.
\end{falsificationbox}

\section{Coq Anchor}

\coqbox{lucas\_2\_eq\_3}{trinity-clara/proofs/igla/lucas\_closure\_gf16.v}{Proven}

\section{Acknowledgement of Provisional Status}

\begin{admittedbox}
This entire chapter is provisional.  It is included so that the spine of
\texttt{main.tex} resolves cleanly during Lane L0 editorial repair; it is
\emph{not} part of the defended text until Lane L33 closes.
\end{admittedbox}
